---
title: "[ECo 602 - Analysis of Environmental Data](/eco_602_634_2020/index.html){target='_blank'}"
subtitle: "Dual Models"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    # toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(cowplot)
```

```{r graphics setup, include=FALSE}

```


# Learning Objectives



# Key Concepts

- Sample spaces
- Common distributions
- Fitting deterministic and stochastic models



# Meet our representative distributions!

- There are hundreds of named *parametric* distributions.  We'll work with two of the most common, and useful, in this assignment:
  - The Normal distribution
  - The Binomial distribution
- These two distributions have many complementary properties that will help you learn how to apply the concepts to other distributions.





## The Normal distribution






## The Binomial distribution



# Model fitting


## Choosing a determinsitic model


## Choosing a stochastic model


## Assessing model fit




# Instructions




# Lab Questions

