---
title: "Using distribution functions"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

```{r, echo = FALSE, fig.cap="Binomial Probabiltiy Mass Function"}
x = 0:5

fill_1 = rgb(0, 0.3, 0.8, 0.25)
fill_0 = rgb(0, 0.0, 0.0, 0.0)

barplot(
  dbinom(x, size = 5, prob = 0.75), 
  col = c(fill_0, fill_0, fill_0, fill_0, fill_1, fill_1), 
  space = 0, names.arg = 0:5,
  xlab = "success count")
```



<div class="questionborder">
### Q1: 3 pts

- What is the probability of observing *more than* 3 successes in a binomial distribution with parameters n = **5** and p = 0.75?
- Paste the R code that you used to find your answer.

</div>


Solution
========

0.632

This is the correct code:
```{r}
1 - pbinom(q = 3, size = 5, p = 0.75)
```

This also works, but it's not generalizable:
```{r}
dbinom(x = 4, size = 5, p = 0.75) +
dbinom(x = 5, size = 5, p = 0.75)
```






Meta-information
================
extitle: Using distribution functions

extype: string

exsolution: nil

exname: Frequentist concepts

exsection: Frequentist concepts

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 3

