---
title: "Using distribution functions"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
require(mfn.teaching.utils)
require(ggplot2)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


```{r, echo = FALSE, fig.asp=0.5, fig.cap="Normal Distribution - Probability Density Function"}
fill_lower = rgb(0, 0, 0, 0)
fill_upper = rgb(0, 0.3, 0.8, 0.25)
fill_middle = rgb(0, 0, 0, 0)
mfn.teaching.utils::plot_norm_tails(lower_tail = NULL, upper_tail = pnorm(1.2, mean = 2, sd = 2), fill_lower = fill_lower, fill_upper = fill_upper, fill_middle = fill_middle, pop_mean = 2, pop_sd = 2) +
  scale_x_continuous(breaks = c(-4, -2, 0, 1.2, 2, 4, 6))
```



<div class="questionborder">
### Q1: 3 pts.

- What is the probability of observing a value of *greater than* 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?
- Paste the R code that you used to find your answer.

</div>


Solution
========

0.655

This is the code they should have used:

```{r}
1 - pnorm(1.2, mean = 2, sd = 2)
```






Meta-information
================

extitle: Using distribution functions

extype: string

exsolution: nil

exname: Frequentist concepts

exsection: Frequentist concepts

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 3

