---
title: "Sampling Distribution"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
require(mfn.teaching.utils)
require(ggplot2)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

```{r, echo = FALSE, fig.asp=0.5}
```

### Central Limit Theorem

Navigate to the <a href="https://seeing-theory.brown.edu/probability-distributions/index.html">probability theory</a> page on Seeing Theory.

Follow the link to the **Central Limit Theorm** section.


### Sampling Distribution

1. Choose a set of $\alpha$ and $\beta$ parameters that result in a skewed beta distribution.
1. Set the **sample size** to 15 and the **draws** to 50.
1. Hit the sample button several times and observe the evolution of the histogram.



<div class="questionborder">
### Q1 2 pts.

- Describe how the shape of the histogram changes as you continue to press the *sample* button.


</div>







Solution
========


```{r}

```






Meta-information
================
extitle: Sampling Distribution

extype: string

exsolution: nil

exname: Frequentist concepts

exsection: Frequentist concepts

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 2

