---
title: "Variables 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



Write R code to create the following variables:

1. A variable `a` that contains the text of your first name.
2. A variable `b1` that contains the number 45.6
3. A variable `b2` that contains the text "45.6"
4. A variable `c1` that contains the sequence of integers from 0 to 3




<p class="blackborder">
Hint: use the `class()` function to help with the questions below.
</p>


- Q1 (1pt): What type of data is contained in the variable `a`
- Q2 (1pt): What type of data is contained in the variable `b1`
- Q3 (1pt): What type of data is contained in the variable `b2`
- Q4 (2pt): What happens when you add `b1` and `b2` and why?
- Q5 (1pt): Are the variables `b1` and `c1` of the same type?
- Q6 (3pt): Explain what happens when you add `b1` and `c`.   Consider both the number of elements in each variable and the data type.






Solution
========


Possible answers

```{r}
a = "mike"
b1 = 45.6 
b2 = "45.6" 
c1 = 0:3

class(a)
class(b1)
class(b2)
class(c1)

b1 + b2

# b1 and c1 are different types (numeric and integer), but they are compatible
# so it is possible to add them together.

b1 + c1
```





Meta-information
================
extitle: Variables 1
extype: string
exsolution: nil
expoints: 9
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
