---
title: "Vectors 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Create a vector called `v1` that contains a sequence of integers from -2 to 2.

When you print the contents of `v1`, it should look like this:


```{r, echo = FALSE}
v1 = -2:2
print(v1)
```

Now, use `v1` to create a new vector called `v2` whose elements are the elements of `v1` multiplied by `3`.  It should look like this:

```{r echo = FALSE}
v2 = 3 * v1
print(v2)
```

Finally, calculate the sum of all the elements in `v2`.


Paste the following lines of code into the text input window below:

- Q1 (1pt): The R code you used to create `v1`.
- Q2 (1pt): The R code you used to create `v2`.
- Q3 (1pt): The R code you used to calculate the sum of elements in `v2`.


Solution
========


Possible answers

```{r eval = FALSE}
v1 = -2:2
print(v1)

v2 = 3 * v1
print(v2)

print(sum(v2))

```





Meta-information
================
extitle: Vectors 1
extype: string
exsolution: nil
expoints: 3
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
