---
title: "Creating a list"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```

Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Create a list, named `my_list_1` with following three elements:

1. first element is numeric:  `5.2`
2. second element is a string `"five point two"`
3. third element is a vector of all integers from 0 to 5 [how do you do this?]

Name the elements in `my_list_1`:

1. "two"
2. "one"
3. "three"

<div class="warning"> Make sure the elements in your list are in the order specified.</div>


- Q1 (2 pts): Paste the R code you used to create `my_list_1` into the text entry box below.


Solution
========


Some possible answers

```{r, echo = TRUE, results='hide'}

# using the names() function
my_list_1 = list(5.2, "five point two", 0:5)
names(my_list_1) = c("two", "one", "three")

# creating the names when you build the list
my_list_1 = list(two = 5.2, one = "five point two", three = 0:5)
```

Other approaches are fine too, as long as they produce the same answer.

The following should return TRUE

```{r}
ref_answer = list(two = 5.2, one = "five point two", three = 0:5)
identical(ref_answer, my_list_1)
```






Meta-information
================
extitle: Creating a list
extype: string
exsolution: nil
expoints: 2
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
