---
title: "Subsetting a list 2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

You will use the `my_list_1` object that you created in the previous question.


<p class="blackborder">
Hint: remember the subsetting operators `[]` and `$`?
</p>


Paste the R code for the the following questions into the text input box below:

- Q1 (1pt): The third element.
- Q2 (1pt): The element with the name "one".  Note: there are at least two ways to do this!



Solution
========

Some possible answers are:

```{r, eval=FALSE, echo = TRUE}
my_list_1[[3]]
my_list_1$one
my_list_1[["one"]]

```


Meta-information
================
extitle: Subsetting a list 1
extype: string
exsolution: nil
expoints: 2
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
