---
title: "Logical tests 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---

```{r echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Run the following code to build a vector called `my_vec`:

```{r}
my_vec = rep(1:3, 5)
my_vec
```

Use the logical equality test operator `==` to create a vector, `my_bool_vec`, of Boolean values from `my_vec`.  

- `my_bool_vec` should be the same length as `my-vec`.
- `my_bool_vec` should have `TRUE` values in the positions where `my_vec` has values of `3`.

```{r echo=FALSE, results='hide'}
my_vec = rep(1:3, 5)
my_bool_vec = my_vec == 3
data.frame(my_vec, my_bool_vec)
```
You can run the following code to check that you have the correct values in `my_bool_vec`:

```{r}
data.frame(my_vec, my_bool_vec)
```

- Q1 (3 pts.) Paste the R code that you used to create `my_bool_vec` into the text entry window.


Solution
========

A possible solution:

```{r, echo=FALSE}
my_vec = rep(1:3, 5)
my_bool_vec = my_vec == 3
data.frame(my_vec, my_bool_vec)
```


Meta-information
================
extitle: Logical tests 1
extype: string
exsolution: nil
expoints: 3
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
