---
title: "Logical subsetting 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Note: You will use `my_vec` and `my_bool_vec` from the previous question.

Use `my_bool_vec` to retrieve all of the elements of `my_vec` that have a value of `3`.


<p class="blackborder">
Hint: Use the square bracket subsetting operator: `[]`.
</p>


```{r echo = FALSE, results='hide'}
my_vec = rep(1:3, 5)
my_bool_vec = my_vec == 3
```

Your result should look like this:

```{r, echo = FALSE}
my_vec[my_bool_vec]
```

- Q1 (2 pts.) Paste the R code that you used to perform the subsetting into the text window below.


Solution
========

A possible solution:

```{r, echo = FALSE}
my_vec[my_bool_vec]
```


Meta-information
================
extitle: Logical subsetting 1
extype: string
exsolution: nil
expoints: 2
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
