---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Week 10 and 11 Reading Questions"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
options(knitr.duplicate.label = "allow")
require(rmd.utils)
require(mfn.teaching.utils)
```



```{r build_assignment_page, eval=FALSE, include=FALSE}
require(rmd.utils)
require(here)

assn_base_name = "reading_questions_week_10"
assn_source_file = find_file("reading_questions_week_10.Rmd", extension = ".Rmd", exact_match = TRUE)
assn_source_dir = dirname(assn_source_file)

question_source_files = find_file(
  "", search_path = file.path(assn_source_dir, "moodle"), 
  return_all = TRUE, extension = ".Rmd")

assn_dir_out    = here::here("docs", "assignments", "eco_602", "reading_questions")
assn_file_out = file.path(assn_dir_out, assn_base_name)

file.exists(assn_source_file)
file.exists(assn_source_dir)
file.exists(assn_dir_out)
file.exists(assn_file_out)

key_file_out = here::here(
  "docs", "answer_keys",
  sprintf("individual_assignment_answer_key_%s", assn_base_name))

build_assignment_doc(
  assn_source_file,
  question_source_files,
  doc_target_file = assn_file_out,
  key_target_file = NULL)

if (FALSE)
{
  build_moodle_xml_questions(
    question_source_files = question_source_files,
    assignment_name = assn_base_name,
    xml_output_path = here::here("docs", "moodle_quiz_questions"))
}


```



# Readings

The reading questions this week are drawn from material in the readings:

- Bolker 9.3, 9.4
- Zuur 6, 7.1 - 7.4



## Reading Guide

# Acronyms

Unfortunately, these readings (and the regression world in general) require you to recognize the differences among may similar sets of names and acronyms.

## The Two Gs

You'll encounter two important G-words:

- General
- Generalized

*General* applies only to the Group 1 methods.  Generalized applies to several of the Group 2 methods.



## Grouping by Assumption 

- Do you remember the the 4 key assumptions?

Each of the various groups of regression models was designed to accommodate one or more violations of these key assumptions.

For each of the groups of models, I suggest you keep a list of which of which violations are permissible. This will help you remember what the different model classes are, and also remind you of situations in which each group would be helpful.


# Questions
