---
title: "Bolker 1.4: Ratio of predation proportions"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---



```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

> The Bolker reading was difficult....

<div class="questionborder">

### Q1 3 pts.

- Use the seed predation proportions you calculated to determine the ratios of seed proportions.

Note: Use the predation rate for *Pseudospondias microcarpa* in the denominator.  Also note that predation proportions (predation rates) are different than odds ratios.

</div>



Solution
========




```{r answer_key, results='hold'}
rm(list = ls())

pol_n_predation = 26
pol_n_no_predation = 184
pol_n_total = pol_n_predation + pol_n_no_predation
pol_predation_rate = pol_n_predation / pol_n_total

psd_n_predation = 25
psd_n_no_predation = 706
psd_n_total = psd_n_predation + psd_n_no_predation
psd_predation_rate = psd_n_predation / psd_n_total


psd_predation_rate / pol_predation_rate
pol_predation_rate / psd_predation_rate
```




Meta-information
================
extitle: Ratio of predation proportions
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 3

