---
title: "Library of Babel: word length"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---



```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


> "Those examples allowed a librarian of genius to discover the fundamental law of the Library. This philosopher observed that all books, however different from one another they might be, consist of identical elements: the space, the period, the comma, and the twenty-two letters of the alphabet. ..."

- A set of 25 characters doesn't seem unmanageable.

> "He also posited a fact which all travelers have since confirmed: In all the Library, there are no two identical books. From those incontrovertible premises, the librarian deduced that the Library is "total"-perfect, complete, and whole-and that its bookshelves contain all possible combinations of the twenty-two orthographic symbols (a number which, though unimaginably vast, is not infinite)-that is, all that is able to be expressed, in every language. ..."


- That sounds like a permutations/combinations problem.



<div class="questionborder">

- There are 25 possible *words* consisting of a single character in the Library.
- There are $25 \times 25 = 25^2 = 625$ possible 2-character words.

### Q1 2 pts

- How many 3-character words are possible?

</div>



Solution
========

Three spaces: $25^3 = 15625$




Meta-information
================
extitle: Library of Babel - word length
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 8
expoints: 2

