---
title: "Library of Babel: library size"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---



```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


NOTE: I wanted students to get 25 * B, but most did not get it. 

I need to do a better lead up next year


> There are $25^{1,312,000}$ possible books in the Library.



- Each book in the library has the same number of available positions for the 25 possible characters:

$410 \times 40 \times 80 = 1,312,000$

- Let's define the variable $B$ as the number of books in the Library of Babel:

<div class="questionborder">
### Q1 2 pts

- How many books would the Library contain if you added *one* additional position to the book size?
  - You shouldn't write out any numbers, provide you answer in terms of B.
  
</div>



Solution
========


The expanded library would contain 25 * B books.



Meta-information
================
extitle: Library of Babel - library size
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 5
expoints: 2

