---
title: "Random Normal Data 3"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---

```{r, echo = FALSE, results = "hide", message = FALSE}
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

> Consider the figure with three histograms you created.


<div class="questionborder">
### Questions

- Q1 (1 pt) Qualitatively describe the differences in the patterns you observe.
- Q2 (2 pts) Explain why they histograms are different.
</div>


Solution
========

Students should realize that as the sample size increases, the histograms are smoother and smoother, approximating a normal curve.





Meta-information
================
extitle: Random Normal Data 3
extype: string
exsolution: nil
exname: lab 4
exsection: lab 4
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
expoints: 3
