---
title: "Simulated Data 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide", message=FALSE} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Instructions:



1. Review the section of the lab instructions in which I created a `data.frame` of random data.
```{r, echo = TRUE, eval=FALSE}
n_pts = 10
x_min = 1
x_max = 10
x = runif(n = n_pts, min = x_min, max = x_max)
dat = data.frame(x = x, y_observed = rnorm(n_pts))
```
2. Experiment with creating and plotting random data:
  - Try different numbers of points.
  - Try different plotting characters.
  - Try using `set.seed()` with different seed values.
3. Create scatterplots of four of your favorite random data sets.
  - Arrange your scatterplots in a single figure: it should be a 2 by 2 grid of scatterplot panels.
4. Save your figure to a file called `sim_data_scatterplots.png`


<div class="questionborder">
### Questions

- Q1 (3 pts.) Paste the R code you used to create one of the random datasets in your figure into the text entry box.
- Q2 (3 pts.) Paste the R code you used to create your figure into the text entry box.
- Q3 (1 pt.) Paste your figure file into the file entry box.

</div>




Solution
========

```{r}
set.seed(12345)
n_pts = 100
x_min = 1
x_max = 10
x = runif(n = n_pts, min = x_min, max = x_max)
dat = data.frame(x = x, y_observed = rnorm(n_pts))

par(mfrow = c(2, 2))
plot(dat$x, dat$y_observed, pch = 23, col = 1, bg = 22)
plot(dat$x, dat$y_observed, pch = 21, col = 1, bg = "red")
plot(dat$x, dat$y_observed, pch = 21, col = 1, bg = "red", lwd = 3)
plot(dat$x, dat$y_observed, pch = 24, col = 1, bg = "red")

```





Meta-information
================
extitle: Simulated Data 1
extype: string
exsolution: nil
exname: lab 4
exsection: lab 4
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 7
