---
title: "Bolker 1.4: Null Hypothesis"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

> The Bolker reading was difficult....

Bolker used a seed predation experiment to illustrate the statistical frameworks.

The primary question in his examples is: *Do seed predation rates vary among* species?


A ***null hypothesis*** is used in the Frequentist paradigm as a baseline against which to compare data.


<div class="questionborder">

### Q1 2 pts

- Describe a possible null hypothesis about the seed predation rates. Paste your answer into the text entry box below.

</div>




Solution
========

The most obvious null is: "There is no difference in seed predation rates among species."


Meta-information
================
extitle: null hypotheses
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 2
