---
title: "Bolker 1.4: Seed Predation Rates"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

> The Bolker reading was difficult....

I have found that working through some of the calculations in a difficult reading helps me to understand what the author meant.

In that spirit, you'll be using the data presented in Bolker Table 1.2 (on the top of page 11) to complete the template code and calculate the seed predation rates.


### Code Template:

```{r template_code, echo=TRUE, eval=FALSE}

# Clear your R environment to make sure there are no stray variables.
rm(list = ls())


pol_n_predation = 26
pol_n_no_predation = 184
pol_n_total = ????
pol_predation_rate = ????

psd_n_predation = ????
psd_n_no_predation = ????
psd_n_total = ????
psd_predation_rate = ????
```

### Self-test

Run the following code after you have made your calculations.  Your rates should match the *observed proportions* in the Bolker text on page 11.


```{r print_proportions, echo = TRUE, eval=FALSE}
print(
  paste0(
    "The seed predation rate for Polyscias fulva is: ",
    round(pol_predation_rate, digits = 3))) 

print(
  paste0(
    "The seed predation rate for Pseudospondias microcarpa is: ",
    round(psd_predation_rate, digits = 3)))
```

<div class="hintborder">

### Hint

To make sure that your code is written correctly, you need to run it in a empty R environment.

The call `rm(list = ls())` on the first line of the code template removes all variables from the environment. Make sure you include it in your code!
</div>

<div class="questionborder">

### Q1 5 pts.

- Paste your completion of the code in the code template above into the text entry field below.

</div>


Solution
========




```{r answer_key, results='hold'}
rm(list = ls())

pol_n_predation = 26
pol_n_no_predation = 184
pol_n_total = pol_n_predation + pol_n_no_predation
pol_predation_rate = pol_n_predation / pol_n_total

psd_n_predation = 25
psd_n_no_predation = 706
psd_n_total = psd_n_predation + psd_n_no_predation
psd_predation_rate = psd_n_predation / psd_n_total

print(
  paste0(
    "The seed predation rate for Polyscias fulva is: ",
    round(pol_predation_rate, digits = 3))) 

print(
  paste0(
    "The seed predation rate for Pseudospondias microcarpa is: ",
    round(psd_predation_rate, digits = 3)))

```




Meta-information
================
extitle: Bolker 1.4: Seed Predation Rates
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 5
