---
title: "Parametric and non-parametric inference"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="questionborder">

### Q1 2 pts.

- Describe a major difference between inference using a *parametric distribution* and *non parametric* methods.

</div>



Solution
========


- The main difference is that parametric assumes the data can be modeled by a parametric distribution, while nonparametric has not distribution assumption.
  - Although... it is usually assumed that the distributions of values in both populations are similar (even if they don't conform to any parametric distribution)

- Students may also notice that nonparametric is usually less powerful.


```{r answer_key, results='hold'}
rm(list = ls())

pol_n_predation = 26
pol_n_no_predation = 184
pol_n_total = pol_n_predation + pol_n_no_predation
pol_predation_rate = pol_n_predation / pol_n_total

psd_n_predation = 25
psd_n_no_predation = 706
psd_n_total = psd_n_predation + psd_n_no_predation
psd_predation_rate = psd_n_predation / psd_n_total


psd_predation_rate / pol_predation_rate
pol_predation_rate / psd_predation_rate
```




Meta-information
================
extitle: Parametric and non-parametric inference
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 2

