---
title: "Two-Way ANOVA: Graphical Exploration"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}

cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructions">

### Instructions

- Examine the boxplot of penguin body mass conditioned on both sex and species:

```{r sex_species_boxplot, echo=FALSE}
bp = 
  boxplot(
  body_mass_g ~ sex + species, 
  data = palmerpenguins::penguins, 
  ylab = "Body Mass (g)",
  las = 2, xlab = "", plot = FALSE)


  boxplot(
  body_mass_g ~ sex * species, 
  data = palmerpenguins::penguins, 
  ylab = "Body Mass (g)",
  las = 2, xlab = "", 
  names = gsub("\\.", "\n", bp$names))


# par(mar = c(11, 4, 0, 0))
# boxplot(
#   body_mass_g ~ sex + species, 
#   data = palmerpenguins::penguins, 
#   ylab = "Body Mass (g)",
#   las = 2, xlab = "")
```

- Compare to the boxplot conditioned only on species:

```{r species_boxplot, echo=FALSE}
boxplot(
  body_mass_g ~ species, 
  data = palmerpenguins::penguins, 
  ylab = "Body Mass (g)", las = 2, xlab = "")

```

</div>


<div class="questions">

### Q 2 pt (short answer)

- Do you think adding sex to a model that already includes species will improve the model fit?
  - Make sure you justify your answer based on the boxplots and not results of a statistical test.

</div>


Solution
========

### Grading Guide

Students should notice that the widths of the IQR in the boxplots are much narrower in the 2-factor plots than the species-only plot.


Meta-information
================

extitle: Two-Way ANOVA: Graphical Exploration 

extype: string
exsolution: nil

expoints: 2

exname: individual_week_10
exsection: individual_week_10

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 12



