---
title: "Two-Way ANOVA: Build the Model"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructions">

### Instructions

- Fit a [factorial] linear model of penguin body mass predicted by penguin sex and penguin species.
  - Be sure to include the sex predictor before the species predictor.
- Save the output of `lm()` to a variable called `fit_both`.

</div>
<div class="warnings">
Warning: in this case we need to fit a 2-way ANOVA model *with interactions*.  You have to use the star symbol instead of the plus symbol to make sure that R includes all of the factor combinations.
</div>

<div class="questions">

### Q 1 pt (R code)

- Paste the R-code you used to build `fit_both`.

<div class="hints"> NOTE: your code should be only one line </div>


</div>




Solution
========

The key is that they need to build a *factorial* model using the star and not the plus symbols.

```{r q3_model_fit}
fit_both = lm(body_mass_g ~ sex * species, data = palmerpenguins::penguins)
summary(fit_both)
```



Meta-information
================

extitle: Two-Way ANOVA: Build the Model
extype: string
exsolution: nil

expoints: 1

exname: individual_week_10
exsection: individual_week_10

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
