---
title: "Two-Way ANOVA: Build the Model"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructions">

### Instructions

When you fit a model with two categorical predictors, the *base case* is now a combination of the base level of predictor one and the base level of predictor two.

- Use `summary(fit_both)` to examine the model coefficient table.

</div>


<div class="questions">

### Q 1 pt (Short answer)

- What is the base case for the two-way model that includes sex and species?

</div>




Solution
========

The base case is female Adelie penguins.

```{r}
aggregate(body_mass_g ~ sex * species, data = palmerpenguins::penguins, FUN = mean, na.rm = TRUE)

fit_both = lm(body_mass_g ~ sex * species, data = palmerpenguins::penguins)
summary(fit_both)

```




Meta-information
================

extitle: Two-Way ANOVA: Build the Model
extype: string
exsolution: nil

expoints: 1

exname: individual_week_10
exsection: individual_week_10

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
