---
title: "Model Coefficients 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructions">

### Instructions

> Recall that you can use the model coefficients to calculate group means when you have a model with categorical predictors.

- Use `summary(fit_both)` to examine the model coefficient table.

</div>

<div class="questions">

### Q 1 pt (Short answer)

- What are the names of the two coefficients (from the first column of the coefficient table) you need to calculate the average mass of female Chinstrap penguins?

</div>




Solution
========

You need:

- Intercept
- speciesChinstrap

The mass is: 3368.8 + 158.4 = 3527.2

If students used an additive model they will get slightly different answers:

3372.39 + 26.92 = 3399.31




```{r}
require(penguins)

fit_both = lm(body_mass_g ~ sex * species, data = penguins)
summary(fit_both)

aggregate(body_mass_g ~ sex * species, data = palmerpenguins::penguins, FUN = mean, na.rm = TRUE)


fit_both_additive = lm(body_mass_g ~ sex + species, data = penguins)
summary(fit_both_additive)

```


Meta-information
================

extitle: Model Coefficients 1
extype: string
exsolution: nil

expoints: 1

exname: individual_week_10
exsection: individual_week_10

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 12
