---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Hypothesis Testing 1: A Parametric Frequentist Approach"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = "allow")
require(rmd.utils)
require(ggplot2)
require(mfn.teaching.utils)

```

## Key Concepts and Terms

- Null Hypotheses
- Alternative Hypotheses
- Decision Criterion
- 1- and 2-tailed hypotheses
- T-test: applying the decision criterion




## A Classical Approach

### Hypothesis testing is central to Frequentist statistics.

- Frequentist hypothesis testing often uses *parametric* distributions, and their 
accompanying assumptions to do inference.


- This lecture uses a t-test example.  T-tests are easy to understand in 
a Frequentist parametric paradigm!

\vfill
\vfill

### Other approaches to hypothesis testing

Note: there are other possible approaches that use Frequentist concepts, such
as simulation and resampling techniques.




```{r child = find_file("hypothesis_testing_frequentist_t_test.Rmd")} 
```



## Key Concepts and Terms

- Null Hypotheses
- Alternative Hypotheses
- Decision Criterion
- 1- and 2-tailed hypotheses
- T-test: applying the decision criterion

