---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Confidence Intervals I: Background"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rmd.utils)

options(knitr.duplicate.label = "allow")

source_file_1 = 
  find_file("confidence_intervals_background.Rmd", search_path = here::here("course_docs", "lectures", "modules"))
  
source_file_2 = 
  find_file("confidence_intervals_calculating.Rmd", search_path = here::here("course_docs", "lectures", "modules"))

```




## Key Concepts and Terms

- Population distribution
- Sample distribution
- Sampling distribution
- Standard error
- [Hypothetical] repeated sampling



## Outline

### Confidence intervals aren't that interesting...

- This slide deck focuses on building up to CIs from the underlying *sampling distribution*
- If you understand the sampling distribution, you've got the tools to understand much of frequentist inference (not just the confidence interval part).




```{r child = source_file_1}
```


## Key Concepts and Terms

- Population distribution
- Sample distribution
- Sampling distribution
- Standard error
- [Hypothetical] repeated sampling





