--- 
title: "Weekly Topic Overview"
output: html_document
editor_options:
  chunk_output_type: inline
--- 

```{r, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
  } else if (knitr::is_html_output()) {
    cat(sprintf("<span style='color: %s;'>%s</span>", color, x))
  }
}
wed_start = as.Date("2020- 08- 26")
dat_fmt = "%b %d"
dat_str = function() return(paste0(format(wed_start, dat_fmt)))
```

```{r increment_date, include = FALSE, eval=FALSE}
wed_start = wed_start + 7
```



# Weekly Lab Schedule



 
## Week 1: `r dat_str()` - Course Software
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```

Assignments:

- Learning Objectives Assessment
- Lab 1: Course software, 



## Week 2: `r dat_str()` - How to R Like a Programmer
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```
 
- Week 3: `r dat_str()` - Data Exploration and Deterministic Functions 1
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```

 
- Week 4: `r dat_str()` - Data Exploration and Deterministic Functions 2
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```


 
- Week 5: `r dat_str()` - Random Numbers and Distributions
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```


 
- Week 6: `r dat_str()` - Dispersion, Noise, and Uncertainty
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```


 
- Week 7: `r dat_str()` - Quantifying Error
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```


 
- Week 8: `r dat_str()` - Linear Models 1
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```


 
- Week 9: `r dat_str()` - Linear Models 2
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```


 
- Week 10: `r dat_str()` - Linear Models 3
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```

 
- Week 11: `r dat_str()` - Simulation
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```

- Week 12: `r dat_str()` - Multivariate
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```

 
- Week 13: `r dat_str()` - Wrap Up

