---
title: "ECo 634 - Analysis of Environmental Data Lab"
author: "Michael France Nelson"
subtitle: "Lab Schedule and Concept Summaries"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    # toc: TRUE
    # toc_float: TRUE  
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
wed_start = as.Date("2020-08-26")
dat_fmt = "%b %d"
dat_str = function() return(paste0(format(wed_start, dat_fmt)))



```


```{r increment_date, include = FALSE, eval=FALSE}
wed_start = wed_start + 7
```


# Lab Assignment Schedule


<details><summary> Lab 1: R Fundamentals 1 `r dat_str()` </summary> 

Key Concepts

- Operators
- Variables
- Comments
- Built-in functions
- Custom functions
- Filesystems

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```



<details><summary> Lab 2: R Fundamentals 2 `r dat_str()` </summary>

Key Concepts

- Data representation in R
- High- and low-level programming
- Chaining functions
- Variable scope
- Importing data
- Basic plots

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```



<details><summary> Lab 3: Data Explortion `r dat_str()` </summary>

Key Concepts

- Data exploration: 
- Deterministic functions: visual parameter estimation

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 4: Uncertainty and Error `r dat_str()` </summary>

Key Concepts

- Pseudo random numbers
- Probability distributions
- Quantifying error
- Minimizing error

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 5: Uncertainty, Samples, and Populations`r dat_str()` </summary>

Key Concepts

- Null hypotheses
- Simulating repeated samples
- Categorical data in R

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 6: Intro to Inference, Confidence Intervals, and Significance`r dat_str()` </summary>

Key Concepts

- T-tests
- Non-parametric tests
- Significance
- False negatives
- False positivies

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```



<details><summary> Lab 7: Least Squares, Categorical Data `r dat_str()` </summary>

Key Concepts

- Model objects in R
- ANOVA in R
- Model coefficient tables
- ANOVA tables


</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 8: Modeling Continuous Data 1 `r dat_str()` </summary>

Key Concepts

- Fitting models
- Simple and multiple regression with `lm()`
- Graphical and numerical model summaries

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 9: Modeling Continuous Data 2 `r dat_str()` </summary>

Key Concepts

- Modeling continuous and categorical data
- Model diagnostics
- Model selection

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 10: Extending the Linear Model `r dat_str()` </summary>

Key Concepts

- What is linearity?
- Fitting a deterministic model
- Logistic regression
- Nonlinear least squares

</details>

```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 11: Frameworks `r dat_str()` </summary>

Key Concepts

- Optimization: Maximum Likelihood
- Bayesian models
- What even is statistics?

</details>



<details><summary> Lab 12: Extending our toolkit `r dat_str()` </summary>

Key Concepts

- Loops
- Custom functions
- The Tidyverse
 
</details>


```{r ref.label = "increment_date", include = FALSE, echo = FALSE}
```


<details><summary> Lab 13: A Prieview of What's Next `r dat_str()` </summary>

Key Concepts

- Unmet assumptions: dealint with difficult data
- Multivariate methods
- Spatial data
- Random walks

</details>

