---
title: "Heterogeneity"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructions">

Consider the various types of models and data transformations from the readings and lecture notes.

</div>


<div class="questions">

### Q 2 pts (short answer)

- Name two possible solutions to deal with heterogeneity, i.e. non-constant variance, in your data.

</div>



Solution
========

## Grading Guide

Answers include:

- data transformation
- Using a Group 2 technique, for example:
  - Using a GLM
  - Using an additive model
  - Weighted least squares


Meta-information
================

extitle: Heterogeneity

extype: string
exsolution: nil

exname: week_10_reading_questions
exsection: week_10_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10

expoints: 2
