---
title: "Base Case"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructions">

Consider an experiment looking at plant biomass response to water treatments.

- The three water level treatments are: "low", "med", and "high". 
- The response is plant biomass in grams after 7 weeks of growth.

|                 | Estimate| Std. Error| t value| Pr(>&#124;t&#124;)|
|:----------------|--------:|----------:|-------:|------------------:|
|(Intercept)      | 2.4     |       2.19|  98.371|              0.001|
|waterMed         | 1.3     |       5.12|   0.480|              0.231|
|waterHigh        | 13.6    |       3.48|  24.495|              0.001|

</div>


<div class="questions">

### Q 1 pt (short answer)

- What is the *base case* for water treatment?

</div>





Solution
========

### Grading Guide

Low is the base case.

From the model coefficient table you can see that "low" is missing.


Meta-information
================

extitle: Base Case

extype: string
exsolution: nil

exname: week_10_reading_questions
exsection: week_10_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5

expoints: 1
