---
title: "Model Coefficients"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructions">

Consider an experiment looking at plant biomass response to water treatments.

- The three water level treatments are: "low", "med", and "high". 
- The response is plant biomass in grams after 7 weeks of growth.

|                 | Estimate| Std. Error| t value| Pr(>&#124;t&#124;)|
|:----------------|--------:|----------:|-------:|------------------:|
|(Intercept)      | 2.4     |       2.19|  98.371|              0.001|
|waterMed         | 1.3     |       5.12|   0.480|              0.231|
|waterHigh        | 13.6    |       3.48|  24.495|              0.001|

</div>


<div class="questions">

### Q 1 pt (numeric entry)

- What was the average plant mass, in grams, for the **low** water treatment?
</div>





Solution
========

### Grading Guide

Answer is 2.4 

Students could do the calculation, or realize that low water is the base case and the coefficient is given as the intercept in the table.

```{r solution}
coeffs = c(2.4, 1.3, 13.6)
x      = c(  1,   0,    0)

sum(coeffs * x)
```


Meta-information
================

extitle: Model Coefficients

extype: num
exsolution: 2.4
extol: 0.01

exname: week_10_reading_questions
exsection: week_10_reading_questions

expoints: 1
