---
title: "Model Summary"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructions">


Consider an experiment looking at plant biomass response to water treatments.

- The three water level treatments are: "low", "med", and "high". 
- The response is plant biomass in grams after 7 weeks of growth.

|                 | Estimate| Std. Error| t value| Pr(>&#124;t&#124;)|
|:----------------|--------:|----------:|-------:|------------------:|
|(Intercept)      | 2.4     |       2.19|  98.371|              0.001|
|waterMed         | 1.3     |       5.12|   0.480|              0.231|
|waterHigh        | 13.6    |       3.48|  24.495|              0.001|

</div>

<div class="questions">

### Q 3 pt (multiple choice)

Which of the following questions *cannot* be addressed with the **model coefficient table**?


Select all the correct answers.

Answerlist
----------
* Is there a positive relationship between increased water availability and plant biomass accumulation?
* Is water availability a significant predictor for plant biomass accumulation?
* What is the average biomass of plants in the high water treatment?

</div>


Solution
========

### Grading guide

* incorrect: you can observe the sign and significance of the water coefficient.
* correct: you need the ANOVA table to determine this.
* incorrect: you can use the coefficients to build the model equation that to calculate this value.


Answerlist
----------
* incorrect: you can observe the water slope coefficient
* correct: you need the ANOVA table to determine this
* incorrect: you can use the coefficients to build the model equation that to calculate this value.



Meta-information
================

extitle: Model Summary

extype: mchoice
exsolution: 010
exshuffle: 3

exname: week_10_reading_questions
exsection: week_10_reading_questions

expoints: 3
