---
title: "significant model coefficients"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructions">

Consider this table of model coefficients from a plant growth experiment with the following *continuous* predictor variables.

Note: The amount of water, N, and P given to each plant was randomized at the beginning of the experiment.

- Water: 3 - 30 mL per week
- Nitrogen: 1.1 - 42 mg per week
- Phosphorus: 1.1 - 42 mg per week


The response variable was plant biomass accumulation after 17 weeks (in grams).

|                 | Estimate| Std. Error| t value| Pr(>&#124;t&#124;)|
|:----------------|--------:|----------:|-------:|------------------:|
|(Intercept)      | -1.7    |       0.23|  98.371|              0.061|
|water            |  0.043  |      0.001|   0.480|              0.021|
|nitro            |  0.192  |      0.034|   1.495|              0.007|
|phosph           | -0.027  |      0.014|   0.091|              0.721|

</div>




<div class="questions">

### Q 2 pt (multiple choice)

- Which of the following predictor variables had slope coefficients that were significantly different from zero at a 95% confidence level?

Select all correct answers:

Answerlist
----------
* water
* nitrogen
* phosphorus

</div>



Solution
========

### Grading guide

Water and Nitgoren both have $p \lt 0.05$, Phosphorus does not.


Answerlist
----------
* correct: p-value less than 0.05
* correct: p-value less than 0.05
* incorrect: p-value greater than 0.05



Meta-information
================

extitle: significant model coefficients

exname: week_10_reading_questions
exsection: week_10_reading_questions

extype: mchoice
exsolution: 110
exshuffle: 3

expoints: 2
