---
title: "Model Coefficients"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructions">

Consider this table of model coefficients from a plant growth experiment with the following *continuous* predictor variables.

Note: The amount of water, N, and P given to each plant was randomized at the beginning of the experiment.

- Water: 3 - 30 mL per week
- Nitrogen: 1.1 - 42 mg per week
- Phosphorus: 1.1 - 42 mg per week


The response variable was plant biomass accumulation after 17 weeks (in grams).

|                 | Estimate| Std. Error| t value| Pr(>&#124;t&#124;)|
|:----------------|--------:|----------:|-------:|------------------:|
|(Intercept)      | -1.7    |       0.23|  98.371|              0.061|
|water            |  0.043  |      0.001|   0.480|              0.021|
|nitro            |  0.192  |      0.034|   1.495|              0.007|
|phosph           | -0.027  |      0.014|   0.091|              0.721|

</div>


<div class="questions">

### Q 1 pt (numeric input)

- Using the information in the model coefficient table above, calculate the expected biomass for a plant given:

  - 0 mL water per week
  - 0 mg nitrogen per week
  - 0 mg phosphorus per week


</div>



Solution
========

### Grading Guide

The answer is -1.7

Students can read the intercept value directly form the coefficient table.

Alternatively they could use dummy variable values with the model coefficients.

```{r}

coeffs = c(-1.7, 0.043, 0.192, -0.027)
treats = c(1, 0, 0, 0)

sum(coeffs * treats)

```




Meta-information
================

extitle: significant model coefficients

exname: week_10_reading_questions
exsection: week_10_reading_questions

extype: num
exsolution: `r fmt(sum(coeffs * treats), 3)`
extol: 0.05


expoints: 1
