---
title: "model types"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```

Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructions">

Recall the constellation of methods and the types of response variables that each can accommodate.

</div>

<div class="questions">

### Q 1 pt (multiple choice)

- Which of the following model types could use for a study that aims to **predict** bird **presence/absence** by percentage of old growth forest?

Select all correct answers:

Answerlist
----------
* Simple linear regression
* Poisson regression
* Logistic regression
* 1-way Analysis of Variance

</div>


Solution
========

### Grading guide

* wrong - Simple linear regression: Doesn't make sense for binary outcomes.
* wrong - Poisson regression: Doesn't make sense for binary outcomes.
* right - Logistic regression: LR works for binary outcomes, like presence/absence
* wrong - 1-way Analysis of Variance: ANOVA doesn't do prediction.




Answerlist
----------
* wrong - Simple linear regression: Doesn't make sense for binary outcomes.
* wrong - Poisson regression: Doesn't make sense for binary outcomes.
* right - Logistic regression: LR works for binary outcomes, like presence/absence
* wrong - 1-way Analysis of Variance: ANOVA doesn't do prediction.



Meta-information
================

extitle: model types

exname: week_10_reading_questions
exsection: week_10_reading_questions

extype: mchoice
exsolution: 0010
exshuffle: 5

expoints: 1
