---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Data Scales"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    # theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)

source(here::here("rmd_tools", "find_file.r"))
source(find_file("proportional_image_crop.R"))

source(list.files(path = here::here("formatting"), pattern = "rmd_functions.R", recursive = TRUE, full.names = TRUE))

# source(find_file("ggplot_confints.R"))

# require(ggplot2)
# require(cowplot)

```



## Key Concepts

- What do we mean by *scale*?
- Measurement scale terms:
  - Discrete, continuous
  - Numeric, categorical
  - Ratio, interval



  
  
  
## Scale

The word *scale* has a lot of meanings... In our context *measurement scale* or *data scale* refers to:

- A measurement scheme that answers questions like
  - Is it numerical?
    - Can it have negative numbers?
    - Can can it contain fractions?
    - Is there a *true zero*?
  - Is it categorical?
    - Is there a meaningful ordering of the categories?



## Scale

A measurement scale is what we use to *quantify* a variable, i.e. an *attribute* of a sampling unit.

### The choice of scale may be context-dependent:

- Does it reflect an intrinsic property of a variable?
- Does it reflect an intrinsic property of how we choose to measure it?

For example: age measured in years vs. age measured in age class.



## Numeric or Categorical

### Is our variable *quantitative* or *qualitative*?

Numeric measurement scales: our variable is measured as a numeric quantity.

Qualitative: our variable can be classified into a *category*.

### Does our *qualitative* variable have a sensible *order*?

- Categories that have a meaningful order are called *ordinal*
  - There may be an order, but the inter-category intervals cannot be directly compared.
- If there is not an intrinsic order they are *nominal*.




## Numeric Scales: Discrete and Continuous

### Discrete cannot take on intermediate values
- Counts, presence/absence, etc

### Continuous variables can *in principle* take on any intermediate value:
- They are *real* or *rational* numbers.
- Our ability to measure may not capture intermediate values, but they are still continuous.




## Interval and Ratio scales

### Is there a *true zero*?

- Degrees in Kelvin: *absolute zero* is the absence of movement of particles.  
  - You can't go lower than *absolute zero*.
- Degrees in Celsius: zero is centered around the freezing point of water.
  - Negative values are possible.
  
### Interval scales can have negative numbers
### Ratio scales are non negative



## Circular Scales

### Circular scales wrap a maximum value back to zero

- Circular scales are not as common, but they occur when thinking about angles.

- Examples include wind direction and aspect
  - Both are measured in degrees (or radians).

### Modulo operator
- The wrapping behavior of circular scales is just like addition in *modular algebra*.



## Circular Scale: Aspect

An example of circular data in practice: the aspect of ungulate herd location observations in mountainous terrain:

```{r, echo=FALSE, out.width='7.7cm', fig.align='center'}
proportional_image_crop("envelope_simulation_aspect.png", height_pct = 0.95, width_pct = 0.87, height_offset = 0.05, width_offset = 0.01)
```









## Converting Among Scales

### Sometimes it's convenient, or necessary, to convert a variable to a different scale.

For example:  consider count data that consist of mostly 0 and 1, with only a few values greater than 1.  

- It might be useful to convert this to binary data, i.e. presence/absence.

### Aggregating into categories

- Age, size, weight classes
- These convert numeric into ordinal scales.



## Converting Among Scales

### Some conversions are *destructive*: you lose information in the conversion process.

- Aggregating numeric scales into categories.
  - e.g. converting absolute ages into age classes:
    - If you know the age, you can assign to an age class.  
    - If you know only the age class, you can't make the reverse transformation.




## Data Scales and Statistical Models

### Most of the time we build models using numbers, on either discrete (integers) or continuous scales.

### Theoretical (parametric) distributions as models

- Remember that we fit models to data, not the other way around.
- Sometimes theoretical distributions are only *approximately* good models for our data.
  - For example: the Normal distribution often fits data, such as weights, very well, but its sample space includes negative numbers.
  - Some discrete distributions that are useful for count data allow for unrealistically large observations.




## Recap

- What do we mean by *scale*?
- Measurement scale terms:
  - Discrete, continuous
  - Numeric, categorical
  - ratio, interval

