---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "R and RStudio"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(cowplot)

g_tit = ggtitle("Motor Trend Car Road Tests")
wt_lab = "Weight (lbs * 1000)"
mpg_lab = "Miles Per Gallon"
hp_lab = "Gross Horsepower"
cyl_lab = "Number of Cylinders"
gear_lab = "Number of Front Gears"

t_bot = theme(legend.position = "bottom")


gg_mod = ggplot(mtcars, aes(mpg, disp)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  ylab("Displacement (cu. in)") +
  xlab(mpg_lab)
```


## Key concepts

- What is R?
- Why should I learn it?
- What is RSTudio?






## What is R

R is a statistical *programming language*.
R has many desireable qualities, for example:

- It is open-source and extensible
- There are thousands of *packages* for just about any kind of analysis you might need to do.
- R is *somewhat easy* to learn compared to many other languages.
  - It is a *high level* language: many technicalities occur behind the scenes so code is compact.
  - It is [mostly] *vectorized*
  - It is an *interpreted* language, so more forgiving of errors than a *compiled* language.




## Why should I learn R

Some good reasons:

- Reproducibility, Collaboration, and Communication
- I'm enrolled in this course
- Graphics capabilities
- Tons of learning materials
- Increasingly a standard tool for research, government, and industry
- Dominant platform in many fields
  - Huge user knowledge base





## Why should I learn R

Did I mention that it's free and open source?





## Why shouldn't I learn R?

Point-and-click programs can be easier to learn:

- Excel and other spreadsheets can handle some calculations and analyses
- It's too hard to learn
  - It's *hard* to learn, but *too hard*.... I'm not convinced.
  - Any programming language has a steep learning curve
- It’s not established in my field
- [Somewhat] idiomatic syntax






## What is RStudio

RStudio is an Integreated Development Environment (IDE)

- Provides a user-friendlier interface to R
- Provides lots of tools to organize projects
- RSTudio and kinttr + RMarkdown
  - Markdown/markup language for creating webpages, pdfs, presentations and other documents
  - Embed R code and graphics




## Where to get R and RSTudio

R is maintained by the Comprehensive R Archive Network:

- https://cran.r-project.org/


RSTudio is at https://rstudio.com/

- The basic RStudio desktop is free and open-source!







## R examples: Basic plotting 1

```{r, echo = FALSE}
data("mtcars")

plot(
  mtcars$wt, mtcars$mpg,
  main = "Motor Trend Car Road Tests",
  xlab = "Weight (* 1000)",
  ylab = "Miles Per Gallon")

```

## R examples: Basic plotting2

```{r, echo = FALSE}
data("mtcars")
par(mfrow = c(2, 2))
plot(
  mtcars$wt, mtcars$mpg,
  main = "Motor Trend Car Road Tests",
  xlab = "Weight (* 1000)",
  ylab = "Miles Per Gallon")
plot(
  mtcars$hp, mtcars$wt,
  main = "Motor Trend Car Road Tests",
  ylab = "Weight (* 1000)",
  xlab = "Gross Horsepower")
boxplot(
  wt ~ cyl, data = mtcars,
  main = "Motor Trend Car Road Tests",
  ylab = "Weight (* 1000)",
  xlab = "Number of Cylinders")
boxplot(
  wt ~ gear, data = mtcars,
  main = "Motor Trend Car Road Tests",
  ylab = "Weight (* 1000)",
  xlab = "Number of Forward Gears")

```








## R examples: Fancier plotting (ggplot2 package)


```{r fancy_plot_ggplot, echo = FALSE}


mtcars$cyl = factor(mtcars$cyl)
mtcars$`No. carbeurators` = factor(mtcars$carb)
mtcars$`Engine` = factor(mtcars$vs, labels = c("v-shape", "straight"))
mtcars$`Gears` = factor(mtcars$gear)
# , labels = c("v-shape", "straight"))

plot_grid(
ggplot(mtcars) + 
  geom_point(aes(wt, mpg, colour = cyl), data = mtcars) +
  ylab(mpg_lab) + xlab(wt_lab) + g_tit + t_bot,

ggplot(mtcars) + 
  geom_point(aes(hp, wt, colour = `Engine`)) +
  ylab(wt_lab) + xlab(hp_lab) + g_tit + t_bot,

ggplot(mtcars) + 
  geom_boxplot(aes(
    x = cyl, 
    y = wt, 
    fill = cyl), show.legend = FALSE) +
  ylab(wt_lab) + xlab(cyl_lab) + g_tit,
ggplot(mtcars) + 
  geom_boxplot(aes(
    x = `Gears`, 
    y = wt, 
    fill = `Gears`), show.legend = FALSE) +
  ylab(wt_lab) + xlab(gear_lab) + g_tit,
nrow = 2
)



```




## R examples: Linear Regression

```{r}
fit_1 = lm(disp ~ mpg, data = mtcars)
```


\begin{center}
```{r, echo=FALSE, comment = "", results='asis'}
table.string <- texreg::texreg(fit_1, return.string = TRUE, caption = "Miles per gallon and displacement")
cat(table.string)

# summary(fit_1)
```
\end{center}



## R examples: Linear Regression
```{r, echo=FALSE, message=FALSE}
gg_mod
```

