---
title: "Identify a question or problem"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```

Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Describe the key ecological, economic, aesthetic, or other problems your group identified related to your scenario.

Describe the questions that guided your group's model building.




Solution
========





Meta-information
================
extitle: Identify a question or problem
extype: string
exsolution: nil
exname: in-class model thinking
exsection: in-class model thinking
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
