---
title: "Random Normal Data 2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---

```{r, echo = FALSE, results = "hide", message = FALSE} 
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructionsborder">

Instructions:

1. Create a figure including histograms of your vectors `norm_17`, `norm_30`, and `norm_300`.
- The three histograms should all be on the same figure, arranged horizontally (1 rows, 3 column).
- Each histogram must have an informative title that indicates how many randomly generated data points were used to build the histogram.

2. Save your figure to a file called `lab_04_hist_01.png`.
- Your figure should be 700 pixels high, and 1500 pixels wide.
- It should have a resolution of 180 dpi.


</div>

<div class="questionborder">
### Q1 2 pts

- Paste the R code you used to create your figure into the text entry box.

### Q2 1 pt

- Upload your figure image file into the file upload box.

</div>

<div class="hintborder">
- You'll need to check out the help entry for the appropriate arguments to `png()` to set the width, height, and resolution.
- If you get a strange gray border around your image, try setting the background color argument to transparent: `bg = "transparent"`. 
</div>




Solution
========

```{r q2_solution, results='hold', fig.asp=1/3}
mean_1 = 10.4
sd_1 = 2.4

norm_17 = rnorm(n = 17, mean = mean_1, sd = sd_1)
norm_30 = rnorm(n = 30, mean = mean_1, sd = sd_1)
norm_300 = rnorm(n = 300, mean = mean_1, sd = sd_1)

if (FALSE)
  png(
    file.path(
      find_file(
        filename = "lab_04_uncertainty_and_error", directory = TRUE), 
      "lab_4_hist_1.png"), 
    width = 1500, height = 700, res = 180)

par(mfrow = c(1, 3))
hist(norm_17,  main =  "17 normally-distributed numbers")
hist(norm_30,  main =  "30 normally-distributed numbers")
hist(norm_300, main = "300 normally-distributed numbers")

if (FALSE)
  dev.off()

```





Meta-information
================
extitle: Random Normal Data 1
extype: string
exsolution: nil
exname: lab 4
exsection: lab 4
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 3
