---
title: "Residuals 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Instructions:

1. Use the dataset you chose for the previous question.
2. Using the code in the lab walkthrough, create a column of predicted y-values.
3. Using the code in the lab walkthrough, create a column of residuals.


<div class="questionborder">


The wording of these questions was confusing for the students.


### Questions

- Q1 (2 pts.) Paste the R code you used to create create the columns of predicted values and residuals.
</div>


Solution
========

- Build random data, visually fit a model
```{r}
set.seed(12345)
n_pts = 100
x_min = 1
x_max = 10
x = runif(n = n_pts, min = x_min, max = x_max)
dat = data.frame(x = x, y_observed = rnorm(n_pts))
guess_x = 0
guess_y = -.3
guess_slope = 0.1
```

- Add predicted values and residuals
```{r}
dat$y_predicted = 
  line_point_slope(dat$x, guess_x, guess_y, guess_slope)
dat$resids = dat$y_predicted - dat$y_observed
head(dat)
```




Meta-information
================
extitle: Residuals 1
extype: string
exsolution: nil
exname: lab 4
exsection: lab 4
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
expoints: 2
