---
title: "Random Normal Data 4"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide", message = FALSE} 
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


> Remember the code I used to create a plot of the density function of a *standard normal* distribution:


```{r}
# Generate a vector of x-values
x = seq(-3, 3, length.out = 1000)
y = dnorm(x)

plot(x, y, main = "Standard Normal PDF", type = "l")
abline(h = 0)
```

<div class="instructionsborder">


NOTE: Need to specify that the plot frame has to include the bell part of tue curve


Instructions

1. Use the code above as a template to plot a density curve for a normal distribution with:
- mean = 10.4
- standard deviation = 2.4
2. Include an informative title that states the mean and standard deviation values you used.
3. Save your figure to a file called `norm_1.png`.


</div>

<div class="hintborder">
- Hint: check out the help entry for `dnorm()` for a description of the arguments.
</div>

<div class="questionborder">

### Q1 1 pt 

- What are the parameters and their values for the *standard Normal* distribution?


### Q2 2 pts

- Paste the R code you used to create and save your figure into the text-entry window.


### Q1 1 pts

- Paste the image file of your figure into the file entry box. 
</div>


Solution
========

- The standard normal has mean = 0 and sd = 1.

```{r echo=TRUE, results='hold'}

# Mean and sd for normal density curve:
mean_1 = 10.4
sd_1 = 2.4

# Generate a vector of x-values
x_lim = c(mean_1 - 3 * sd_1, mean_1 + 3 * sd_1)

x = seq(x_lim[1], x_lim[2], length.out = 1000)
y = dnorm(x, mean = mean_1, sd = sd_1)

if (FALSE)
  png(
    file.path(
      find_file(
        filename = "lab_04_uncertainty_and_error", 
        directory = TRUE),
      "norm_1.png"), 
    width = 1200, height = 1400, res = 220)

plot(
  x, y, 
  main = paste0("Normal PDF: mean = ", mean_1, " sd = ", sd_1),
  type = "l")
abline(h = 0)

if (FALSE)
  dev.off()

```

```{r, echo=FALSE, eval=FALSE}

mean_1 = 10.4
sd_1 = 2.4

norm_17 = rnorm(n = 17, mean = mean_1, sd = sd_1)
norm_30 = rnorm(n = 30, mean = mean_1, sd = sd_1)
norm_300 = rnorm(n = 300, mean = mean_1, sd = sd_1)
```


Meta-information
================
extitle: Random Normal Data 4
extype: string
exsolution: nil
exname: lab 4
exsection: lab 4
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 10
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 4
