---
title: "Random Normal Data 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r echo = FALSE, include=FALSE, message = FALSE} 
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructionsborder">


Instructions:

1. Create three vectors of normally-distributed random numbers, `norm_17`, `norm_30`, and `norm_300`.  
  - You should tell R that you want your random deviates have a *mean* of 10.4 and a *standard deviation* of 2.4.
  - `norm_17` should have 17 elements, `norm_30` should have 30 elements, and `norm_300` should have 300 elements.

</div>

<div class="questionborder">

### Q1 1pt

- Paste the R code you used to create your vectors into the text entry box.

</div>

<div class="hintborder">

- Hint: You should save the *parameters* you use for *mean* and *standard deviation* into *variables* so that they are easy to re-use (and easy to change later, if needed).

- NOTE: you need to paste your code from a text editor (such as the script window in RStudio).
  - If you paste your code from the R Console, each line starts with a **`>`** character and it won't run directly when pasted into the console (i.e. when we grade your code).

This is OK:

`norm_17 = ....your code here...`

This is incorrect:

`> norm_17 = ....your code here...`

</div>






Solution
========


```{r answer_key, echo = TRUE, eval = FALSE}
mean_1 = 10.4
sd_1 = 2.4

norm_17  = rnorm(n = 17,  mean = mean_1, sd = sd_1)
norm_30  = rnorm(n = 30,  mean = mean_1, sd = sd_1)
norm_300 = rnorm(n = 300, mean = mean_1, sd = sd_1)
```





Meta-information
================
extitle: Random Normal Data 1
extype: string
exsolution: nil
exname: lab 4
exsection: lab 4
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 1
