---
title: "Badness of fit measure"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



In Chapter 6, McGarigal writes:

> The squared measure of deviation has an accelerating penalty: a deviation that is twice as large
contributes four times as much to the sums of squares. There is no a priori reason to choose
such a measure.


<div class="questionborder">

### Q1 5 pts.

- Propose a possible alternative measure of deviation.  
  - Briefly describe how your measure of deviation could affect the parameter estimates.  
  - You may want to consider things such as the effect of large deviations, and the sign of individual deviations.

</div>


```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```







Solution
========



Meta-information
================

extitle: Badness of fit measure

extype: string

exsolution: nil

exname: week_07_reading_questions

exsection: week_07_reading_questions

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 5

