---
title: "Monte Carlo Randomization"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```


Question
========


```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



In Chapter 6, McGarigal writes:


> With the bootstrap, the original data structure is maintained; that is, the individual observation vectors are left intact.

> With the Monte Carlo randomization procedure, the original data structure is destroyed by randomly shuffling some of the data




<div class="questionborder">

### Q1 2 pts

- Regarding the Monte Carlo randomization, what does McGarigal mean by "... the original data structure is destroyed by randomly shuffling..."?

</div>


Solution
========



Meta-information
================

extitle: Monte Carlo Randomization

extype: string

exsolution: nil

exname: week_07_reading_questions

exsection: week_07_reading_questions

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 2

