---
title: "parametric and non parametric"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```




Question
========


```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Refer back to sections 6.1 and 7.2 for McGarigal's descriptions of the form of the linear statistical model for the non parametric OLS and parametric likelihood-based inference techniques.

Recall that he used the same data to illustrate both paradigms: Brown creeper abundance (response) and proportion of late successional forest (predictor).

<div class="hintborder">
Note: McGarigal specifies the parametric model using this notation:

$Y \sim Normal(a + bx, \sigma)$

However, both the parametric and non parametric model can be expressed in the more familiar regression model format:

$y_i = \beta_0 + \beta_1 x_i + e_i$

</div>



<div class="questionborder">


### Q1 3 pts

Describe the key difference between the non parametric model (Ch. 6.1) and the parametric model (Ch. 7.1)

</div>







Solution
========



Meta-information
================

extitle: parametric and non parametric models

extype: string

exsolution: nil

exname: week_07_reading_questions

exsection: week_07_reading_questions

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 3

