---
title: "interpolation and extrapolation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


McGarigal writes:

- There are two kinds of prediction, and these are subject to very different levels of uncertainty:
  - Interpolation – which is prediction within the measured range of the data.
  - Extrapolation – which is prediction beyond the measured range of the data.



<div class="questionborder">

### Q1 2 pts

- Explain why extrapolation has more potential pitfalls than interpolation.

</div>



Solution
========



Meta-information
================

extitle: interpolation and extrapolation

extype: string

exsolution: nil

exname: week_07_reading_questions

exsection: week_07_reading_questions

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 2

