---
title: "Model comparison"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)

```


Question
========


<div class="instructions">

McGarigal writes:

> "We expect a model with more parameters to fit better in the sense that the negative log-likelihood should be smaller if we add more terms to the model. But we also expect that adding more parameters to a model leads to increasing difficulty of interpretation."

</div>


<div class="questions">

### Q1: 4 pts.

- Consider the trade off between model complexity and interpretability.  

- Describe the inherent conflict between using a complicated model that minimizes the unexplained variation and using a simple model that is easy to communicate.

- Since your answer is targeted to a non-scientist audience, I suggest you use a narrative style using a concrete example.

</div>


```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```







Solution
========



Meta-information
================

extitle: Model comparison

extype: string

exsolution: nil

exname: week_07_reading_questions

exsection: week_07_reading_questions

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 4

