---
title: "Histogram of slope"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Examine the histogram for sampling site **slope**.


<div class="questionborder">

### Q1 1 pt 

- Upload an image file of your histogram.

### Q2 3 pts 

- Interpret the histogram in non-technical language that a non-scientist audience would understand.  Some points to consider:
  - Are most sample sites flat?
  - Is there an even mixture of steep and shallow slopes?

</div>


Solution
========

```{r slope_hist}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

hist(dat_all$slope, xlab = "slope (percent)", main = "Sample Site Slope")
```



Meta-information
================
extitle: Histogram of slope
extype: string
exsolution: nil
expoints: 4
exname: data exploration det. funs
exsection: data exploration det. funs
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
