---
title: "Histogram of aspect"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Examine the histogram for sampling site **aspect**.



<div class="questionborder">

### Q1 1 pt 

- Upload an image file of your histogram.


### Q2 2 pts 

- What is aspect, and how is it measured?


### Q3 3 pts 

- Interpret the histogram in non-technical language that a non-scientist audience would understand.  Some points to consider:
  - Do the sampling sites tend to be on north-facing slopes? 
  - South-facing?
  - Evenly distributed?

</div>

Solution
========


If they don't set a custom breaks option, the students may notice
that the rightmost bin is very short.

- It's due to an artifact of the data format and R-s binning algorithm
  - degrees are in the interval 0 to 360
  - The last bin contains the interval 350 - 400

```{r aspect_hist}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

hist(dat_all$aspect, xlab = "aspect (degrees)", main = "Sample Site Aspect")
```







Meta-information
================
extitle: Histogram of aspect
extype: string
exsolution: nil
exname: data exploration det. funs
exsection: data exploration det. funs
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 6
