---
title: "Terrain and basal area scatterplots"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="questionborder">

### Q1 3 pts

- Upload image files of the three scatterplots you created of the terrain variables and basal area.

</div>


Solution
========

```{r terrain_scatterplots, fig.asp=3}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

pch = 16
p_col = rgb(0, 0.4, 0.1, 0.2)
{
  par(mfrow = c(3, 1), pch = 16, col = p_col)
  plot(ba.tot ~ elev, data = dat_all, main = "Basal area and elevation", ylab = "Total Basal Area")
  plot(ba.tot ~ slope, data = dat_all, main = "Basal area and slope", ylab = "Total Basal Area")
  plot(ba.tot ~ aspect, data = dat_all, main = "Basal area and aspect", ylab = "Total Basal Area")
}

```

Meta-information
================
extitle: Terrain and basal area scatterplots
extype: string
exsolution: nil
exname: data exploration det. funs
exsection: data exploration det. funs
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 3
