---
title: "Terrain and basal area scatterplots: interpretation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Consider the three terrain/basal area scatterplots.


<div class="questionborder">

### Q1 (1 pts).

- Describe the relationship you observe between **elevation** and **basal area**.


### Q2 (1 pts)

- Describe the relationship you observe between **slope** and **basal area**.


### Q3 (1 pts)

- Describe the relationship you observe between **aspect** and **basal area**.


### Q4 (3 pts). 

- Describe how basal area varies with terrain in the sample sites in Oregon.
  - Make sure your answer is in non-technical language that an audience of non scientists would understand.

</div>

Solution
========

```{r terrain_scatterplot_interpretation, fig.asp=3}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

pch = 16
p_col = rgb(0, 0.4, 0.1, 0.2)
{
  par(mfrow = c(3, 1), pch = 16, col = p_col)
  plot(ba.tot ~ elev, data = dat_all, main = "Basal area and elevation", ylab = "Total Basal Area")
  plot(ba.tot ~ slope, data = dat_all, main = "Basal area and slope", ylab = "Total Basal Area")
  plot(ba.tot ~ aspect, data = dat_all, main = "Basal area and aspect", ylab = "Total Basal Area")
}
```


Meta-information
================
extitle: Terrain and basal area scatterplots: interpretation
extype: string
exsolution: nil
exname: data exploration det. funs
exsection: data exploration det. funs
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 6
