---
title: "Logistic model 2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Consider the second bird species you chose to examine.

<div class="questionborder">

### Q1 1pt. 

- Upload an image file of the logistic function plot.

### Q2 3 pts. 

- What was the bird species? Describe its presence/absence patterns in terms of basal area.  Did it seem to prefer areas with high or low tree cover?

</div>
Solution
========

- Student answers will vary.
- A logistic model with a single predictor probably doesn't seem like a great fit for many of these birds.






Meta-information
================
extitle: Logistic function 2
extype: string
exsolution: nil
exname: lab_03
exsection: lab_03
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 4
