---
title: "Counting total sites with jays"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Using the `dat_all` data frame, calculate the **total number of sampling sites** in which Gray Jays were observed.

Hint: What happens when you use the `sum()` function on a vector of Boolean values?


<div class="questionborder">

### Q1 1pt

- Paste the R code you used for the calculation into the text entry window.

</div>



Solution
========

There are 110 stations with gray jays present.



The following works:
```{r, eval=TRUE}
dat_bird = read.csv(here("data", "bird.sta.csv"))
dat_habitat = read.csv(here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)
sum(dat_all$GRJA >= 1)
sum(dat_all$GRJA > 0)
```


Meta-information
================
extitle: Count of presences
extype: string
exsolution: nil
exname: lab_03
exsection: lab_03
exextra[essay,logical]: TRUE
exextra[essay_format,character]: plain
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 1
