---
title: "Likelihood of Two Observations"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructionsborder">

Consider the likelihood calculation example for 2 observations of Wilson's Warblers using a Poisson distribution model.

- Recall the R-code I used in the walkthrough to find the log-likelihood of using $\lambda = 4.5$:

```{r}
wiwa_counts = c(2, 6)
dpois(x = wiwa_counts, lambda = 4.5)
```

- As a group, experiment with different values of $\lambda$ to find a value that maximizes the sum of log-likelihoods of the hypothetical observations.
  - Hint: you only need to consider 3 decimal digits!


</div>

<div class="questionborder">
### 1 pts: 

- What value for $\lambda$ did you select?

</div>





Solution
========

This is the code they should have used:

```{r}
wiwa_counts = c(2, 6)
dpois(x = wiwa_counts, lambda = 4.5)
```

They may have tried some of the following:
```{r}
dbinom(x = 1, size = 4, p = 0.75)
dbinom(x = 2, size = 4, p = 0.75)
dbinom(x = 3, size = 4, p = 0.75)
dbinom(x = 4, size = 4, p = 0.75)
```





Meta-information
================

extitle: Likelihood of Two Observations

extype: string

exsolution: nil

exname: in_class_week_08_1

exsection: in_class_week_08_1

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 1

