---
title: "Binomial Likelihood of Winter Wren Counts"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructionsborder">

In the walkthrough, I tried using a Poisson distribution with $\lambda = 1.0$ to model the set of census counts of Wilson's Warblers.

You will create models for the census count of Winter Wrens.

- Suppose you wanted to use a binomial distribution instead of a Poisson distribution.
- Find values of the parameters for a binomial distribution that make the vector of census counts most likely.


</div>


<div class="questionborder">
### 2 pts:

- What are the two parameters for a binomial distribution and what do they represent?
- What were the parameter values you selected?
- How did you choose a value for $n$?
- Paste the line of R code you used to calculate the binomial log-likelihood for the vector of Winter Wren census counts.

</div>





Solution
========

This is the code they should have used:

```{r}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)



sum(log(dbinom(x=dat_all$WIWR, 9, p=0.18)))

```






Meta-information
================

extitle: Single Value Likelihood

extype: string

exsolution: nil

exname: in_class_week_08_1

exsection: in_class_week_08_1

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 1

