---
title: "Learning Objectives Assessment Q7"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Describe some strategies you have used when you get stuck while working on a problem or debugging code in R (or another language).

Some things to consider:

- R internal help files
- message boards (Stackexchange etc.)
- colleagues, mentors, friends
- have certain resources been more helpful than others?




Solution
========





Meta-information
================
extitle: Learning Objectives Assessment Q7
extype: string
exsolution: nil
exname: ECo 634 Learning Objectives
exsection: ECo 634 Learning Objectives
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
