---
title: "Learning Objectives Assessment Q8"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Explain why you want to learn R and how you think it is a benefit to your academic or research career.

Some possible things to consider:

- Learning curve
- Pros and cons of scripting vs. point-and-click
- use of R in your field



Solution
========





Meta-information
================
extitle: Learning Objectives Assessment Q7
extype: string
exsolution: nil
exname: ECo 634 Learning Objectives
exsection: ECo 634 Learning Objectives
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
