---
title: "Poisson Likelihood of Winter Wren Counts"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructionsborder">

In the walkthrough, I tried using a Poisson distribution with $\lambda = 1.0$ to model the set of census counts of Wilson's Warbler.  

You will create models for the census count of Winter Wrens.

<div class="warn">NOTE: the original wording of the instructions were ambiguous, the following is an update.</div>

- Find the $\lambda$ value of a Poisson model that makes all of the observed the Winter Wren's census counts most likely.

- Plot a histogram of Winter Wren counts.  Make sure you choose an appropriate number of bins for the plot.

- As a group, you'll be searching for the value of $\lambda$ that maximizes the sum of log-likelihood.  In other words, you'll find the the model parameter that makes the values of observed wren census counts most likely.
  - Hint: you only need to consider 2 decimal places in your value of $\lambda$!

</div>


<div class="questionborder">
### 1 pts:

- Paste your histogram into your document.
- What value for $\lambda$ did you select?
- Paste the line of R code you used to calculate the Poisson log-likelihood for the vector of Winter Wren census counts.

</div>





Solution
========

They could calculate the mean value for a good guess for lambda:

```{r}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)


mean(dat_all$WIWR)
sum(log(dpois(x = dat_all$WIWR, lambda = 1.4)))

sum(log(dpois(x = dat_all$WIWR, lambda = mean(dat_all$WIWR))))
sum(log(dpois(x = dat_all$WIWR, lambda = mean(dat_all$WIWR) + 0.01)))
sum(log(dpois(x = dat_all$WIWR, lambda = mean(dat_all$WIWR) - 0.01)))
```

They may have tried some of the following:
```{r}
dbinom(x = 1, size = 4, p = 0.75)
dbinom(x = 2, size = 4, p = 0.75)
dbinom(x = 3, size = 4, p = 0.75)
dbinom(x = 4, size = 4, p = 0.75)
```





Meta-information
================

extitle: Poisson Likelihood of Winter Wren Counts

extype: string

exsolution: nil

exname: in_class_week_08_1

exsection: in_class_week_08_1

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 1

