---
title: "Selecting the Best Model"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="instructionsborder">

You explored the use of two discrete *parametric* distributions to model the Winter Wren census counts.

- Which one was better?

</div>


<div class="questionborder">
### 2 pts: 

- Considering a Maximum Likelihood criterion, which model better fit the data?
- Considering what you know about the Binomial and Poisson distributions, which model is more appropriate for census count data?

</div>





Solution
========

This is the code they should have used:

```{r}
hist(dat_all$WIWR, breaks = 6)


wiwa_counts = c(2, 6)

dpois(x = wiwa_counts, lambda = 4.5)
```

They may have tried some of the following:
```{r}
dbinom(x = 1, size = 4, p = 0.75)
dbinom(x = 2, size = 4, p = 0.75)
dbinom(x = 3, size = 4, p = 0.75)
dbinom(x = 4, size = 4, p = 0.75)
```





Meta-information
================

extitle: Selecting the Best Model

extype: string

exsolution: nil

exname: in_class_week_08_1

exsection: in_class_week_08_1

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 1

