---
title: "homogeneity: penguin body mass and species"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}

cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructions">

#### Instructions

We know that the Group 1 methods require homogeneity (constant variance).

- Review the walkthrough to review how to test for homogeneity of variances.

We conducted Analyses of Variance on the penguin data in the individual assignments in lecture.

- We should verify that the the variances are homogeneous in the groups we specified in our ANOVAs.



Use `bartlett.test()` to check for homogeneity of the variances in our model of penguin body mass as predicted by **penguin species**.


```{r, eval = FALSE, echo = FALSE}
require(palmerpenguins)
boxplot(flipper_length_mm ~ species, data = penguins)
bartlett.test(flipper_length_mm ~ species, data = penguins)
bartlett.test(bill_depth_mm ~ species, data = penguins)
bartlett.test(bill_length_mm ~ species, data = penguins)
bartlett.test(body_mass_g ~ species, data = penguins)

```


</div>

<div class="hints">

- Hint: you will probably want to work through the corresponding material in the lecture assignment before working on this question.

</div>

<div class="questions">

### Q1 1 pt (numeric entry)

- What was the p-value from the Bartlett test of homogeneity for observations grouped by *species*?
  - You can round your answer to 4 decimal digits.

</div>



Solution
========

My implementation:

```{r lab_09_q_1}
require(palmerpenguins)
bartlett.test(body_mass_g ~ species, data = penguins)
```



Meta-information
================
extitle: homogeneity: penguin body mass and species

exname: lab_09
exsection: lab_09


extype: num
exsolution: 0.05005
extol: 0.01

expoints: 1




