---
title: "brown creeper habitat presence/absence"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructions">

#### Instructions

- Review the walkthrough materials on testing proportions and contingency tables.
- Conduct a Chi-square test on the contingency table of Brown Creeper presence/absence in edge and interior habitats.

</div>


<div class="questions">

### Q 2 pt (short answer)

- What is the null hypothesis for your test?
  - Make sure you state the null hypothesis in terms of Brown Creeper presence/absence and edge/interior habitats.

### Q 2 pt (short answer)

- Consider the results of your test and explain whether you think that Brown Creepers show a significant habitat preference.
  - Make sure your use the output of your statistical test to support your answer.

</div>



Solution
========

My implementation:

```{r lab_09_q_1}
birds   = read.csv(here::here("data", "bird.sta.csv"), header=TRUE)
hab     = read.csv(here::here("data", "hab.sta.csv"), header=TRUE)
birdhab = merge(birds, hab, by=c("basin", "sub", "sta"))

# Create a contingency table for edge/interior and brown creeper presence/absence
table(birdhab$s.edge, birdhab$BRCR > 0)

# set the presence to be in the first column
br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]


X = chisq.test(br_creeper_table)

X
str(X)


boxplot()

```   



Meta-information
================
extitle: brown creeper habitat presence/absence

extype: string

exsolution: nil

exname: lab_09
exsection: lab_09

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15


expoints: 4




