---
title: "dispersion power plot"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    includes:
      after_body: !expr here::here("formatting", "js", "rmd.js")
editor_options: 
  chunk_output_type: console
---


```{r solution_code, echo=FALSE}
birds = read.csv(here::here("data", "bird.sub.csv"), header = TRUE)
hab = read.csv(here::here("data", "hab.sub.csv"), header = TRUE)
birdhab = merge(hab, birds, by = c('basin', 'sub'))

fit_1 = lm(BRCR~ls, data = birdhab)
fit_1_summary = summary(fit_1)
sd_obs = fit_1_summary$sigma

dat_dispersion_sim = 
  read.csv(
    here::here(
      "data", "lab_11",
      "sim_output_dispersion_1000_200.csv"))
```
```{r solution_plot, eval=FALSE, echo = FALSE}
plot(
  power ~ sd, data = dat_dispersion_sim, 
  type = "l", 
  xlab = "Population SD",
  ylab = "Statistical Power")
abline(v = sd_obs, col = "red", lty = 3)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructions">

### Instructions

Create a line plot of dispersion (x-axis) and statistical power (y-axis) with a dotted vertical red line at the observed population standard deviation value.

</div>


<div class="questions">

### Q 1 pt (image upload)

- Upload your plot as an image file.

</div>



Solution
========

```{r answer_plot, ref.label = "solution_plot", fig.asp=1/1.8, echo = FALSE}
```

<div class="toggle"><button>Click to show/hide code</button>
<br>

```{r answer_code, ref.label = "solution_code", results='hide'}
```
```{r answer_plot, ref.label = "solution_plot", eval = FALSE}
```

</div>




Meta-information
================
extitle: dispersion power plot

extype: string

exsolution: nil

exname: lab_11
exsection: lab_11

exextra[essay,logical]: TRUE

exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 1

exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE

expoints: 1




<!-- <script> -->
<!--   $(".toggle").click(function() { -->
<!--     $(this).toggleClass("open"); -->
<!--   }); -->
<!-- </script> -->

