---
title: "dispersion n contour plot"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    includes:
      after_body: !expr here::here("formatting", "js", "rmd3.js")
editor_options: 
  chunk_output_type: console
---


```{r solution_code, echo=FALSE}
load(
  file = here::here(
    "data", "lab_11",
    "sim_output_dispersion_n_1000.RData"))
```
```{r solution_plot, eval=FALSE, echo = FALSE}
contour(
  x = sim_3_dat$pop_sd,
  y = sim_3_dat$sample_size, 
  z = sim_3_dat$power,
  xlab = "population s.d.",
  ylab = "sample size",
  main = "Contour Plot of Statistical Power")
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructions">

### Instructions

Create a contour plot of the ***sample size and population dispersion*** power simulation.

</div>


<div class="questions">

### Q 2 pt (image upload)

- Upload your plot as an image file.

</div>



Solution
========

```{r answer_plot, ref.label = "solution_plot", fig.asp=1/1.8, echo = FALSE}
```

<button onclick="toggle()">Click to show/hide code</button>

<div id="toggle">

```{r answer_code, ref.label = "solution_code", results='hide'}
```
```{r answer_plot, ref.label = "solution_plot", eval = FALSE}
```
</div>


Meta-information
================
extitle: dispersion n contour plot

extype: string

exsolution: nil

exname: lab_11
exsection: lab_11

exextra[essay,logical]: TRUE

exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 1

exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE

expoints: 2
