---
title: "dispersion n interactive plot"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    includes:
      after_body: !expr here::here("formatting", "js", "rmd.js")
editor_options:
  chunk_output_type: console
---


```{r setup, include = FALSE}
options(knitr.duplicate.label = "allow")
require(rmd.utils)
```

```{r fun_code, child=find_file("function_chunks_lab_11.Rmd", exact_match = TRUE), include=FALSE}
```
```{r data_code, child=find_file("load_data_chunks_lab_11.Rmd", exact_match = TRUE), include=FALSE}
```
```{r sim_code, child=find_file("simulation_chunks_lab_11.Rmd", exact_match = TRUE), include=FALSE}
```


```{r solution_code, echo=FALSE}
load(
  file = here::here(
    "data", "lab_11",
    "sim_output_dispersion_n_1000.RData"))
```
```{r solution_plot, eval=FALSE, echo=FALSE}
rgl::persp3d(
  x = sim_3_dat$pop_sd,
  y = sim_3_dat$sample_size, 
  z = sim_3_dat$power,
  xlab = "pop. sd",
  ylab = "n",
  zlab = "power",
  col = 'lightblue',
  alpha = 0.7,
  theta = 30, phi = 30, expand = .75,
  ticktype = 'detailed')

```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructions">

### Instructions

Create an interactive 3D perspective plot, using `persp3d` in package `rgl`, of the ***sample size and population dispersion*** power simulation.

You need to save this to an html file using `writeWebGL()` in the package `rgl`.

Experiment with different values for the `width` and `height` arguments to customize the display size of the 3D plot.

  - Your surface must be a different color than <a href="https://michaelfrancenelson.github.io/eco_602_634_2020/webGL/pop_sd_size_power_sim_plot.html" target="_blank"> my example here.</a>contour plot of the ***sample size and population dispersion*** power simulation.

</div>


<div class="questions">

### Q 2 pt (image upload)

Upload your plot as an ***interactive html*** file that you created with writeWebGL().

</div>



Solution
========

```{r answer_plot, ref.label = "solution_plot", fig.asp=1/1.8, echo = FALSE, eval = FALSE}
```

<div class="toggle"><button>Click to show/hide plot code</button>
<br>

```{r answer_code, ref.label = "solution_code", results='hide', eval = FALSE}
```
```{r answer_plot, ref.label = "solution_plot", eval = FALSE}
```
</div>
<div class="toggle"><button>Click to show/hide simulation code</button>
<br>

```{r answer_code, ref.label = "dispersion_n_sim", results='hide', eval=FALSE}
```
</div>


Meta-information
================
extitle: dispersion n interactive plot

extype: string

exsolution: nil

exname: lab_11
exsection: lab_11

exextra[essay,logical]: TRUE

exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 1

exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE

expoints: 2


