---
title: "Formulas and equations"
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "moodle_template_funs.R"))
```



Question
========


```{r CSS, echo = FALSE, results = "asis"}
writeLines(readLines(here::here("formatting", "css", "styles.css")))
```

If the equations:

<div class="blackborder">

> $cov_{x,y}=\frac{\sum_{i=1}^{N}(x_{i}-\bar{x})(y_{i}-\bar{y})}{N-1}$

> $P(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{ - \left( {x - \mu } \right)^2 }/{2\sigma ^2 }}$

</div>
<br>

are intuitive and easy to interpret, great!

If not, describe what seems difficult or opaque about them and how would you approach learning their meanings?


<div class="warn">Note: If you don't know these equations do NOT spend time trying to interpret or understand them now!</div>
<br>
<div class="warn">Note: If they don't make much sense now it is fine!  I don't expect them to seem natural for many of us now. </div>



Solution
========





Meta-information
================
extitle: Formulas and equations 2
extype: string
exsolution: nil
exname: learning objectives assessment
exsection: learning_objectives_assessment
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 8
