---
title: "Formulas and equations"
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "moodle_template_funs.R"))
```


<!-- Fill in the following code to include an image file: -->
```{r, echo = FALSE, results = "hide", eval=FALSE} 
include_supplement(
  "image_file.png",
  dir = here::here(),
  recursive = TRUE)
```




Question
========

```{r CSS, echo = FALSE, results = "asis"}
writeLines(readLines(here::here("formatting", "css", "styles.css")))
```

Describe what **uncertainty** means to you in the context of analysis of environmental data.

Some things you may want to consider:

- Sources of uncertainty
- Quantifying uncertainty
- Uncertainty and statistical significance
- Uncertainty in a frequentist perspective



Solution
========





Meta-information
================
extitle: uncertainty
extype: string
exsolution: nil
exname: learning objectives assessment
exsection: learning_objectives_assessment
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
