---
title: "Type I and II Errors"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Describe statistical type I and type II errors, also known as false positives and false negatives.

Describe when in the experimental design and data analysis process you can specify or calculate the type I and type II error rates.



Solution
========





Meta-information
================
extitle: Type I and II Errors
extype: string
exsolution: nil
exname: learning objectives assessment
exsection: learning_objectives_assessment
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
