---
title: "The Frequentist Paradigm"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Describe the Frequentist paradigm in statistics.

Some things you may want to consider:

- potentially unintuitive ideas in the Frequentist paradigm
- hypothetical infinite sampling
- differences with the Bayesian paradigm
- assumptions
- sample and population



Solution
========





Meta-information
================
extitle: the frequentist paradigm
extype: string
exsolution: nil
exname: learning objectives assessment
exsection: learning_objectives_assessment
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
