---
title: "four assumptions"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



<div class="questionborder">

Zuur lists 4 key assumptions for general linear models.

### Q1: 1 pt. (short answer)

- What are the four assumptions?


</div>




Solution
========

The four assumptions are:

- Independent observations
- Normally-distributed residuals
- Fixed - x
- Constant variance: homogeneity, no heteroskedasticity

Meta-information
================

extitle: four assumptions

extype: string

exsolution: nil

exname: week_08_reading_questions
exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

expoints: 1

