---
title: "choosing a modeling approach"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="questionborder">

> "In the best case, your data will match a classical technique like linear regression
exactly, and the answers provided by classical statistical models will agree with the
results from your likelihood model." - Bolker (2008)

Bolker describes custom-made analyses based on Maximum Likelihood, which often have a biological, ecological, or mechanistic justification.


He contrasts these with the familiar Least Squares, *canned* methods that we typically learn in our first statistics course.

### Q1: 1 pt. (short answer)

- Briefly describe some of the the tradeoffs between using more realistic ML methods and more familiar classical techniques.

</div>




Solution
========

- ML methods are [potentially] harder to code and/or specify
- More complicated models are harder to explain
- Classical statistics/models are well-known and may require less justification
- lots of other possibilities!




Meta-information
================

extitle: choosing a modeling approach

extype: string

exsolution: nil

exname: week_08_reading_questions
exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

expoints: 1

