---
title: "regression and ANOVA"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="questionborder">

### Q1: 1 pt. (short answer)

- Describe a key difference between a simple linear regression and a one-way Analysis of Variance (ANOVA).

<div class="hintborder">
- Hint: consider the data types of the predictor and response variables.
</div>

</div>




Solution
========

My target answer: SLR uses a continuous predictor, one-way ANOVA uses a single categorical variable.

Meta-information
================

extitle: regression and ANOVA

extype: string

exsolution: nil

exname: week_08_reading_questions
exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

expoints: 1

