---
title: "dual-model: regression equation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="questionborder">

We often present the equation for a simple linear regression model as:

$y_i = \alpha + \beta_1 x_i + \epsilon$

### Q1: 1 pt. (short answer)

- Identify the *deterministic* components(s) of the regression equation.

### Q2: 1 pt. (short answer)

- Identify the *stochastic* component(s) of the regression equation.

</div>




Solution
========

Determinsitic: 

- $\alpha$
- $\beta$

Stochastic:

- $\epsilon$


Meta-information
================

extitle: dual-model: regression equation

extype: string

exsolution: nil


exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

expoints: 2

