---
title: "normality"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="questionborder">

> "The normality assumption means that if we repeat the sampling many times
under the same environmental conditions, the observations will be normally distributed
**for each value of X**." - Zuur (2007)

A common misconception about this assumption is that the values of the response variable must be normally distributed.


Consider this histogram of penguin bill lengths:

```{r, echo=FALSE, warning = FALSE, message=FALSE}
require(palmerpenguins)
require(ggplot2)

hist(penguins$bill_length_mm, main = "Histogram of Penguin Bill Length", xlab = "bill length (mm)")
```

- Bill lengths appear very non-normal.
- A the very low p-value in the Shapiro test of normality provides strong evidence against the null hypothesis that bill lengths are normally-distributed.

```{r}
shapiro.test(penguins$bill_length_mm)
```

Nevertheless, a general linear model of bill lengths as that includes species and body mass as predictors passes a test of the normality assumption:

```{r}
fit_1 = lm(bill_length_mm ~ body_mass_g + species, data = penguins)
shapiro.test(residuals(fit_1))
```



### Q1: 3 pt. (short answer/essay)

- Explain how the normality assumption can be met in a general linear model, even if the response variable is not normally-distributed.

</div>




Solution
========


The key is that normality applies to the residuals.  

The response should be normally distributed around the predicted value for each combination of predictor variable values.


Meta-information
================

extitle: normality

extype: string

exsolution: nil

exname: week_08_reading_questions
exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 5

expoints: 3

