---
title: "slope coefficient interpretation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="questionborder">

Consider the regression equation for a simple linear regression:

$y_i = \alpha + \beta_1 x_i + \epsilon$


### Q1: 3 pt. (short answer/essay)

- Describe the meaning of the slope parameter $\beta_1$ in the context of the relationship between the predictor variable, x, and the response variable y.
  - Your answer must be in plain non-technical language.
  - Your explanation will be most effective if you use a narrative approach, using a concrete example to illustrate the concept.

</div>




Solution
========

"A one-unit increase in x will result in a $beta_1$ unit change in y."

A narrative approach using a concrete example is best.



Meta-information
================

extitle: slope coefficient interpretation

extype: string

exsolution: nil

exname: week_08_reading_questions
exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

expoints: 3

