---
title: "Model Selection"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(mfn.teaching.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="questionborder">

> "The first part of the AIC definition is a measure of goodness of fit. The second
part is a penalty for the number of parameters in the model." - Zuur (2007)


### Q1: 3 pt. (short answer/essay)

- Why should a model selection criterion penalize the number of parameters in a model?

</div>




Solution
========

Many possible answers.

Ideally they realize that more complicated models are more difficult to interpret and explain.



Meta-information
================

extitle: model selection

extype: string

exsolution: nil

exname: week_08_reading_questions
exsection: week_08_reading_questions

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 15

expoints: 3

