---
title: "[ECo 602 - Analysis of Environmental Data](/eco_602_634_2020/index.html){target='_blank'}"
subtitle: "Setting up Your ECo 602 Repository"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
require(rmd.utils)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
source(here::here("formatting", "functions", "rmd_functions.R"))
```



# Learning Objectives

- Create a class git repository
- Create a course RProject
- Upload your repository to GitHub


# Class Repository Instructions

## ECo 602 Git Repository

You will be creating a git repository to hold your course R work.  

Instructions:

1. Open up GitHub Desktop and click the appropriate button to **create a new repository on your computer.**

```{r echo = FALSE, results='asis', warning=FALSE}
expandable_html_image("new_repository_github_desktop_2.png", img_width = 60)
```

You will call your repository "ECo_602_2020".

Make sure that you:
- check the box to create a README
- select the **R** option from the **Git Ignore** drop down menu.

<br>

When you're ready, click the button to create the repository.  Make sure you know how to find the folder on your computer that contains the new repository.

<br>

On my computer, the new repository was created here:
```{r results='asis', echo=FALSE, message=FALSE}
expandable_html_image("new_repository_github_desktop_file_location.png")
```

If you look inside the new folder, you should see something like this:

```{r results='asis', echo=FALSE}
expandable_html_image("new_repository_github_desktop_folder_contents.png")
```

The important things to notice are:

1. A directory called *.git*
1. A file called *.gitignore*


<div class="note">
NOTE: Your operating system may hide files and folders that begin with a period character.  You can change the folder view settings to show files and folders that start with a period.

You can find instructions to set your system to show *hidden files* here:

- <a href="https://support.microsoft.com/en-us/help/4028316/windows-view-hidden-files-and-folders-in-windows-10" target="_blank">Windows</a>
- <a href="https://www.businessinsider.com/how-to-view-hidden-files-on-mac" target="_blank">Mac OS</a>
</div>


## Create a RProject

Now you need to create an *RProject* in your new repository directory.

1. Open up RStudio.
2. Select the *New Project* option in the *File* menu.
3. On the popup select the *Existing Directory* option.
```{r echo = FALSE, results='asis'}
expandable_html_image("rstudio_new_project_existing_directory.png")
```
4. Find the location where you created your **ECo_602_2020** repository in GitHub Desktop.
```{r echo = FALSE, results='asis'}
expandable_html_image("rstudio_new_project_select_directory.png")
```

## Course Repository Directory

To make grading, data organization, and collaboration with your classmates, we will all be using a common directory structure for the coding work in this course.

1. In your computer's file browser, find the folder where you created your new repository.  
1. Create the following folders:
  - *data*
  - *individual_assignments*
  - *in_class_assignments*
  - *figures*
  
<br>
Your class code directory should now look like this:
```{r echo=FALSE, results='asis'}
expandable_html_image("602_repo_folder_contents.png", thumb_width = 120, img_width = 40)
```

## Add some files

Now it's time to add some files to your repository:

- Copy the R script you used to do your work for the first lecture R assignment (DataCamp: Intro to R) into the *individual_assignments* folder.

- Copy the R script you used to do your work in the in-class group assignment (Week 2: In-class R) into the *in_class_assignments*.  You should rename the script file to `week_02_group.R`

- Create a new R script called `week_03_data_exploration.R` and save it in the *individual_assignments* folder.

<br>

At this point my individual assignments folder looks like this:

```{r echo=FALSE, results='asis'}
expandable_html_image("602_individual_assignments_folder.png", thumb_width = 120, img_width = 40)
```


## Commit changes and push to GitHub

Now, go back to GitHub Desktop and commit your changes to your repository.  Each time you create a snapshot by committing your changes, you will have to add a short commit message.

You should type the text *First commit!* in the Summary field.

```{r echo=FALSE, results='asis'}
expandable_html_image("github_desktop_ready_to_commit.png", thumb_width = 120)
```

Notice the *Changes* field in the left of the window.  You'll see the names of the new files that you created in your repository directory.

When you're ready, click the **Commit to Master** button in the lower left.

## Push changes to GitHub

Now that you've created a snapshot of your work, it's time to push your changes to github.com.  You should see a button that says **Publish Repository**.  Click this to upload your changes to your github profile.

```{r echo=FALSE, results='asis'}
expandable_html_image("github_desktop_ready_to_push_1.png", thumb_width = 120)
```
```{r echo=FALSE, results='asis'}
expandable_html_image("github_desktop_ready_to_push_2.png", thumb_width = 120)
```


## Check your github profile

Now you should verify that your new repository is up and running on github.com.

Navigate to your profile on github.com and verify that you have a repository called `ECo_602_2020`


## Create a zip file of your course folder.

