---
title: "Install R and RStudio"
---

```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

exams::include_supplement(
  c("install_rstudio_question.PNG",
  "github_desktop_signed_in.PNG"),
  dir = here::here(),
  recursive = TRUE)
```




Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "eco_602_2020.css")))
```

<div class="warn">Before attempting this exercise, follow the instructions for installing R and RStudio in the assignment  description.</div>


# Instructions:

- Start an RStudio session.
- Locate the console window.
- Type the following (replacing the text in quotes with your name) into your console window and press enter:

> "Your Name Here"

- Type the following into the console and press enter:

> getwd()

- Take a screenshot of your RStudio session and upload it in the file upload field.

<div class="blackborder">

Your screenshot should look something like this:

```{r echo = FALSE, results = 'asis'}
expandable_html_image("install_rstudio_question.PNG", moodle_quiz = TRUE, bg_alpha = 0.1)
# expandable_html_image("install_rstudio_question.PNG", moodle_quiz = TRUE, img_width = 70, bg_alpha = 0.1)
```
</div>



- Upload your screenshot in the file upload field below:

Solution
========




Meta-information
================
extitle: Install R and RStudio
extype: string
exsolution: nil
exname: setup_software
exsection: setup_software
exextra[essay,logical]: TRUE
exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
