---
title: "Join DataCamp Course"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide", include = FALSE} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "eco_602_2020.css")))
```

# Instructions

1. Follow the DataCamp course invite link to join the DataCamp course.

2. Select the correct answer from the selection below:

Answerlist
----------
* I have enrolled in the DataCamp Intro to R course.
* I have not enrolled in the DataCamp Intro to R course.




Solution
========

Did you get it right?


Answerlist
----------
* Great!
* You need to follow the invite link on Moodle and join the course on DataCamp.



Meta-information
================
extitle: Join DataCamp Course
extype: schoice
exname: setup_software
exsection: setup_software
exsolution: 10
