---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "EC0 602 Assignment Schedule 2020"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# source(list.files(path = here::here("formatting"), pattern = "rmd_functions.R", recursive = TRUE, full.names = TRUE))
```






## Reading questions

|	Question Set	|	Due Date	|
|	----	|	----	|
|	Reading Questions: Week 2	|	09/02/20	|
|	Reading Questions: Week 3	|	09/15/20	|
|	Reading Questions: Week 4	|	09/18/20	|
|	Reading Questions: Week 5	|	09/25/20	|
|	Reading Questions: Week 6	|	10/02/20	|
|	Reading Questions: Week 7	|	10/09/20	|
|	Reading Questions: Weeks 8/9	|	10/30/20	|
|	Reading Questions: Week 10/11	|	11/06/20	|
|	Reading Questions: Week 12	|	11/13/20	|
|	Reading Questions: Week 13	|	11/20/20	|





## Individual Assignments

|	Name	|	Start Week	|	Due	|
|	----	|	----	|	----	|
|	Learning Objectives Assessment	|	1	|	Fri., Sep 4: Week 2	|
|	Install Course Software	|	1	|	Sep 4: Week 2	|
|	DataCamp: Intro to R	|	1	|	Sep 4: Week 2	|
|	Course GitHub Repo	|	2	|	Fri., Sep 11: Week 3	|
|	Deterministic Functions	|	3	|	Fri, Sep. 18: Week 4	|
|	Dual Models	|	4	|	Fri. Oct 2: Week 6	|
|	Frequentist Concepts	|	6	|	Fri. Oct 9: Week 7	|
|	Using Models 1	|	7	|	Fri. Oct 23: Week 9	|
|	Using Models 2	|	9	|	Fri. Nov 6: Week 11	|
|	Using Models 3	|	11	|	Fri. Nov 20: Week 13	|


<!-- |	Begin Week	| Name	|	Due	| -->
<!-- |	----	|	----	|	----	| -->
<!-- |	1	|	Install Course Software	|	9/4/2020	| -->
<!-- |	1	|	DataCamp: Intro to R	|	9/4/2020	| -->
<!-- |	1	|	Learning Objectives Assessment	|	9/4/2020	| -->
<!-- |	2	|	Course GitHub Repo	|	9/11/2020	| -->
<!-- |	3	|	Deterministic Functions	|	9/18/2020	| -->
<!-- |	4	|	Dual Models	|	10/2/2020	| -->
<!-- |	6	|	Frequentist Concepts	|	10/9/2020	| -->
<!-- |	7	|	Using Models 1	|	10/23/2020	| -->
<!-- |	9	|	Using Models 2	|	11/6/2020	| -->
<!-- |	11	|	Using Models 3	|	11/20/2020	| -->


```{r}
require(palmerpenguins)






aggregate(body_mass_g ~ species + sex, data = penguins, FUN = mean, na.rm = TRUE)
aggregate(body_mass_g ~ species, data = penguins, FUN = mean, na.rm = TRUE)



fit_both2 = lm(body_mass_g ~ species * sex, data = penguins)
fit_both1 = lm(body_mass_g ~ species + factor(sex), data = penguins)

summary(fit_both1)
summary(fit_both2)


boxplot(body_mass_g ~ sex * species, data = penguins)

3368 + 158

3527.206






```


