---
title: "ECo 602 - Analysis of Environmental Data"
author: "Michael France Nelson"
subtitle: "Course Syllabus"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
    # number_sections: TRUE
  # prettydoc::html_pretty:
  #   theme: cayman
  #   highlight: github
  #   css: !expr here::here("formatting", "css", "eco_602_2020.css")
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    toc_float: TRUE
header-includes:
  \input{`r here::here("formatting", "css", "eco_602_2020.tex")`}
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here::here("formatting", "functions", "rmd_functions.R"))
shortcuts_text =
'
<span>
Here are some shortcuts to selected sections:

<ul>
<li><a href="#grading">Grading</a></li>

<li><a href="#course_materials">Course Materials</a></li>
</ul>
</span>

'
```

`r html_output_only(shortcuts_text)`



#  Course Information {.tabset}


## Instructor
Dr Michael F. Nelson

Email: michaelnelso@umass.edu

Phone: 651-308-5430

Direct physical mail to the Environmental Conservation Department office:

> University of Massachusetts
> 160 Holdsworth Way
> Amherst, MA 01003-9285
> 1 (413) 545-2665



## Teaching Assistant

Amanda Suzzi (Ph.D student in ECo)

Email: asuzzi@umass.edu

## Course Meetings

Live course meetings: Tuesdays and Thursdays 8:30 to 9:45 AM

Office Hours: Tuesdays and Thursdays 1:00 - 2:00 PM and by appointment

Course Zoom URL: https://umass-amherst.zoom.us/j/94612508944?pwd=SW5hMmtZWnRYbGtXbW0rd0dwazdqUT09

Zoom Password: Please find the password in the course Moodle.

<!-- Course Zoom Password: **DataIsCool** -->

<div class="info"> NOTE: This course is fully remote, there will be no in-person meetings.  All course meetings, breakout groups, and office hours will be via Zoom. </div>


## UMass Syllabus Guidelines


The following sections are to be included in course syllabi per the UMass faculty senate guidelines.

This syllabus includes information on each of topics in the, indicated syllabus sections, as well as additional sections specific to this course in the body of this syllabus document.




- **Course objecives** are in the **Learning Objectives** and **Course Themes** sections.

- **Expectations and requirements** such as papers, lab reports or exams are in the **Course Structure** and sections.

- **Attendance policies** in the **Course Structure** section.

- **Grading criteria** and the approximate weight of each course requirement in the final grade are in the **Grading** section.

- **Examination schedule** and any **make up or rescheduling policies**. Information on these topics is found in the **Grading**, **Final Take-home Exam**, and **Course Structure** sections.

- Policies on **academic honesty**. Please see the the **Accommodation and Academic Honesty** section.

- **Office, phone and, mailbox numbers** of instructor (instructors should include preferred online contact information (i.e. instructor or course email or use of Blackboard/Moodle communication tools) are in the **Course Information** section.



```{r child = '602/eco_602_course_description.Rmd'}
```

<a name="learning_objectives"></a>
```{r child = '602/eco_602_learning_objectives.Rmd'}
```

<a name="course_materials"></a>
```{r child = '602/eco_602_materials.Rmd'}
```

```{r child = '602/eco_602_course_themes.Rmd'}
```

```{r child = '602/eco_602_course_structure.Rmd'}
```

<a name="grading"></a>
```{r child = '602/eco_602_grading.Rmd'}
```

```{r child = '602/eco_602_weekly_overview.Rmd'}
```

<!-- ```{r child = '602/eco_602_teaching_philosophy.Rmd'} -->
<!-- ``` -->

```{r child = 'accommodation_academic_honesty.Rmd'}
```
