---
title: "flipper bootstrap histogram"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Question
========

```{r CSS, echo = FALSE, results = "asis"}

cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructions">

#### Instructions

- Use the `two.boot()` function to calculate 10000 bootstrap replicates of the difference in mean flipper length of Chinstrap and Adelie penguins.
- Is there missing data?  If so, what argument have we used before in functions like `mean()` and `sd()` to exclude NA values?
- Save the output of `two.boot()` to a variable called `pen_boot`.  This is so that I can replicate your code easily on my machine for grading and assistance.

- Plot a histogram of your bootstrapped differences and save it to an image file.

</div>

<div class="hints">

- Hint: you'll need to investigate the structure of `pen_boot` to figure out what component contains the bootstrapped differences.

</div>

<div class="questions">

### Q1 1 pt. - (file upload)

- Upload your histogram.

</div>



Solution
========

### Grading Guide:

- The histogram should be centered around approx. -6
- SD should be about 1.0


Sample implementation:
```{r lab_8_q_01, fig.width = 4, fig.height = 3, message=FALSE}
require(palmerpenguins)
require(simpleboot)
require(boot)

penguin_dat = droplevels(subset(penguins, species != "Gentoo"))

pen_boot = 
  two.boot(
    subset(penguin_dat, species == "Adelie")$flipper_length_mm,
    subset(penguin_dat, species == "Chinstrap")$flipper_length_mm,
    FUN = mean,
    R = 10000,
    na.rm = TRUE
  )

sd(pen_boot$t)
hist(
  pen_boot$t,
  main = "Bootstrapped differences",
  sub = "Adelie and Chinstgrap Penguins",
  xlab = "Difference in mean flipper length (mm)")

```



Meta-information
================
extitle: flipper bootstrap histogram

extype: string

exsolution: nil

exname: lab_08
exsection: lab_08

exextra[essay,logical]: TRUE

exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 1
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE

expoints: 1



