---
title: "pine seedling Wilcoxon test"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
```

<div class="instructionsborder">

#### Instructions

-  Conduct a Wilcoxon ranked sum test on the difference in the mean number of pine seedlings in between the treatments control and clipped treatments.
  - You should use the *formula* notation in R to conduct the test.

</div>

<div class="questionborder">

### Q1 1 pt. (R code)

- Upload the code you used to conduct the test.

### Q2 1 pt. (numeric answer)

- What was the p-value?

</div>



Solution
========

### Grading Guide:

- Q1: `wilcox.test(pine ~ treatment, dat = dat_tree)`
- Q2: p-value should be 0.1005.

Sample implementation:
```{r q_06_answer_key}
veg = read.csv(here::here("data", "vegdata.csv"), header=TRUE)
dat_tree = droplevels(subset(veg, treatment %in% c("control", "clipped")))
wilcox.test(pine ~ treatment, dat = dat_tree)
```



Meta-information
================
extitle: pine seedling Wilcoxon test

extype: string

exsolution: nil

exname: lab_08
exsection: lab_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 12

expoints: 2



