---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Hypothesis Testing"
author: "Michael France Nelson"
date: "Fall 2020"
output:
    beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    # theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(mfn.teaching.utils)
require(rmd.utils)

require(magick)
require(cowplot)
require(ggplot2)

img_bluegill = image_read(find_file("bluegill.jpg"))
img_bluegill_2 = image_read(find_file("bluegill_dead.jpg"))
g_bluegill = print(ggdraw() + draw_image(img_bluegill))

g = g_bluegill


dev.off()

plot_grid(g, g, g, g, g, g, g, nrow = 1)


ggdraw(clip = "on") + draw_image(image_read(find_file("that_word_does_not_mean_what_you_think_it_means.jpg")))
ggdraw(clip = "off") + draw_image(image_read(find_file("that_word_does_not_mean_what_you_think_it_means.jpg")))


```

```{r, echo = FALSE, fig.height = 3, fig.align='center'}

```



## Frequentist Confidence and Significance

```{r bluegill_banner, echo=FALSE, warn= FALSE,fig.height=1}
g = g_bluegill
plot_grid(g, g, g, g, g, g, g, nrow = 1)
```

### What we want a confidence interval to mean:

- Based on the results of my research, I am 95% sure that the true mean length of bluegill in Massachusetts is between 9 and 10cm.

### What a confidence interval actually means:

- If I could repeat my research procedure many times, I expect that the true mean length of bluegill would fall within my 95% confidence intervals approximately 95% of the time.

```{r bluegill_banner_2, echo=FALSE, warn= FALSE,fig.height=1.5}
i = magick::image_flip(img_bluegill_2)
g = ggdraw() + draw_image(i)
plot_grid(g, g, g, g, g, nrow = 1)
```



## Frequentist Confidence and Significance

### Frequentist statements of confidence and significance do not refer the outcome of a *specific* experiment!

- This isn't exactly what we want from our analysis.
- It's difficult to explain the "if I repeated..." idea to a nonscientific audience.


It's not that a Frequentist analysis doesn't let us make conclusions from an experiment, it's just that we have to be careful to acknowledge the subtly of focusing on the process in the context of repeated sampling, rather than a single outcome.

### Watch for how popular media describes statistical results!




## Recap of Key Concepts and Terms

- What is the Frequentist paradigm?
- Assumptions
- Frequentist pros and cons
- Confidence and significance

