---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Sources of Error"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

smp_error_pois = function(
  n1, lambda, seed = NULL,
  bins = 10,
  mn = NULL, xl = NULL, yl = NULL, 
  mar = c(2, 2.2, 3, 0))
{
  if (!is.null(seed)) set.seed(seed)
  par(mfrow = c(2, 2), mar = mar)
  
  for (i in 1:4)
  {
    x1 = rpois(n1, lambda)
    d1 = density(x1)
    hist(x1, freq = FALSE, main = mn, xlab = xl, ylab = yl, breaks = bins); 
    points(d1$x - 0.5, d1$y, type = "l")
  }
}


```




## Sources of Error

McGarigal describes 3 sources:

- Mesurement
- Process
- Model

I'd like to add a fourth, underappreciated, source of error: **sampling error.**




## Measurement Error

### We cannot measure with perfect precision or accuracy.

How can we commit measurement errors?

- Flawed instruments
- User error
- Failing to observe individuals or species that are actually present
- Over- or under-counting




## Process Error

### There is randomness inherent in any natural system

- Demographic stochasticity
- Environmental stochasticity





## Model Error

\begin{blueborder}{0.9\textwidth}{What if my model is wrong?}
Your model \textit{is} wrong!  
\end{blueborder}


### A mis-specified model can arise from:

- omission of important predictors
- inclsuion of unimportant predictors
- complicated dependencies in predictor variables: multicollinearity.
- mis-specified relationships: linear, exponential, lograthmic
- wrong deterministic model
- incorrect error structure: wrong stochastic model




## Sampling Error

### Sampling error is due to *nonrepresentative* sampling

Sampling is, or should be, a realization of a ***stochastic process***.

- Different realizations produce different samples.
- Some samples are not representative!
- What if you get 10 heads in a row?
- You can reduce, but never eliminate, sampling error by increasing sample size

### Genetic drift is a form of sampling error




## Sampling Error: Graphical Intuition n = 20

```{r echo=FALSE}
n1 = 20
lambda = 6.4
seed = 12345
smp_error_pois(n1, lambda, seed = seed)
```




## Sampling Error: Graphical Intuition n = 40

```{r echo=FALSE}
n1 = 40
lambda = 6.4
seed = 12345
smp_error_pois(n1, lambda, seed = seed)
```




## Sampling Error: Graphical Intuition n = 80

```{r echo=FALSE}
n1 = 80
lambda = 6.4
seed = 12345
smp_error_pois(n1, lambda, seed = seed)
```




## Sampling Error: Graphical Intuition n = 200

```{r echo=FALSE}
n1 = 200
lambda = 6.4
seed = 12345
smp_error_pois(n1, lambda, seed = seed)
```




## Sampling Error: Graphical Intuition n = 1000

```{r echo=FALSE}
n1 = 1000
lambda = 6.4
seed = 12345
smp_error_pois(n1, lambda, seed = seed)
```

