---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Stochastic Processes"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(cowplot)
require(animation)
source(here::here("rmd_tools", "find_file.R"))
source(find_file("rwalk_1.R"))
```



## Stochastic Process

\begin{blueborder}{0.9\textwidth}{What is a stochastic process?}\begin{itemize}\item A procedure whose outcome is uncertain, or random.
\item When we carry out a realization of a stochastic process, we produce an outcome, e.g. a sampling \textit{event}.
\item The range of possible outcomes of a stochastic process is the \textit{sample space}.
\end{itemize}
\end{blueborder}

Probability theory studies stochastic processes

- It's nice when we can associate elements in the sample space with probabilities.






## Some Stochastic Process Examples

### stochastic processes

- Flipping a coin, drawing a hand of cards, lottery draw
- A random walk, Brownian motion


### stochastic process realizations 

- A book in the Library of Babel: Oh tiempo, tus piráides
- A count of Brown Creepers in a forest patch in Oregon




## Random Walks: 50 steps

```{r random_walk_50, echo = FALSE}
n = 50
par(mfrow = c(2, 3), mar = c(0, 0, 0, 0))
for (i in 1:6)
{
  dat = rwalk_1(n)
plot(
  y ~ x, data = dat, 
  type = "l", ann = FALSE, frame = TRUE, axes = FALSE, 
  lwd = 10, col = gray(0, 0.1),
  asp = 1)
points(
  y ~ x, data = dat, 
  type = "l", 
  lwd = 1, col = rgb(0, 0, 0.7, 1),
  asp = 1)
}
```

## Random Walks: 5000 steps

```{r random_walk_5000, echo = FALSE}
n = 5000
par(mfrow = c(2, 3), mar = c(0, 0, 0, 0))
for (i in 1:6)
{
  dat = rwalk_1(n)
plot(
  y ~ x, data = dat, 
  type = "l", ann = FALSE, frame = TRUE, axes = FALSE, 
  lwd = 10, col = gray(0, 0.1),
  asp = 1)
points(
  y ~ x, data = dat, 
  type = "l", 
  lwd = 1, col = rgb(0, 0, 0.7, 1),
  asp = 1)
}
```