---
title: "[ECo 602 - Analysis of Environmental Data](/eco_602_634_2020/index.html){target='_blank'}"
subtitle: "Final Project"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

```


# Creating a successful code example

## Self-Contained, concise, reproducible examples

A successful code example should communicate what the function or operator does.

The best examples are:

- Self-contained:
  - If possible, the example does not rely on external data sets.
  - To the greatest extent possible, the example does not rely on external packages.
    - This isn't always possible, it's ok to include external packages if they contain essential data or supporting functions!
- Concise:
  - The example is as simple and brief as possible.
  - The example focuses on the function or operator in question, avoiding other functions as much as possible.
  - Avoid jargon and technical terminology as much as possible.
- Reproducible:
  - The example must run on anyone's computer!

Your code examples need to be written in your own words, in such a way that you will be able to decipher them later.


## Code Comments

You should use comments in your code as needed to highlight important points.




# Sample Code Example

The following is a code example for the `c()` function.

<div class="instructions">

The function `c()` *combines* or *concatenates* its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

- All of the elements must be of the same *type*.
  - I can't combine character and numeric types in the same call to `c()`

Here's two examples using numeric and character data types:

```{r}
# Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

# Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

I can show the contents of a vector by typing the name of the vector, or using the `print()` function.

```{r}
# Typing the name of the vector into the console prints the contents
num_vec

# The print() function accomplishes the same task:
print(char_vec)
```
</div>





# Data Types and Subsetting

Your 

## Data Structures

- vectors
  - recall that these are often created using `c()`.
- `data.frame()`
- `matrix()`

## Subsetting Operators and Functions

- `[]`
- `$`
- `subset()`



# Required Functions and Arguments

All functions are denoted with parentheses.

The required function arguments are contained in indented bullet points below the corresponding function.


## Functions 1: Data Structure Functions

- `c()`
- `matrix()`
- `data.frame()`
- `length()`
- `nrow()`
- `ncol()`
- `dim()`



## Functions 2: Numerical Data Exploration

- `summary()`
- `mean()`
- `sd()`
 


## Functions 3: Graphical Data Exploration

- `par()` required arguments:
  - `mrow =`
  
- `plot()` required arguments:
  - `col =`
  - `pch =`
  - `cex =`
  - `main =`
  - `xlab =`
  - `ylab =`
  - `xlim =`
  - `ylim =`

- `hist()` required arguments:
  - `breaks =`
  
- `boxplot()`
  - Your must include two examples: a simple boxplot (i.e. one box) and a conditional boxplot (i.e. a single plot that has multiple boxes, conditioned on a categorical variable).
  

## Functions 4: Distribution Functions

- `dnorm()`
- `pnorm()`
- `qnorm()`

- `dbinom()`
- `pbinom()`
- `qbinom()`


## Functions 5: Other Functions

- `subset()`
- `libary()` and `require()`
- `read.csv()`


