---
title: "Weekly Topic Overview"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r, include=FALSE}


colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
  } else if (knitr::is_html_output()) {
    cat(sprintf("<span style='color: %s;'>%s</span>", color, x))
  }
}
```


# Weekly Concepts Overview

This is a brief overview of the topics covered each week.

Please see the detailed weekly schedule for more specific information on readings, assignments, mini-lectures, etc.


```{r, results = 'asis', echo = FALSE}
if (knitr::is_html_output())
{
colorize('Click on  week to expand or hide.\n', "blue")
  # colorize("abc", "blue")
}
```


<!-- # Weekly Concepts Overview {.tabset} -->


<details>
<summary> 
Week 1: August 24, 28 – Introductions, Why Model? 
</summary>

- Introduction to course 
- Course themes 
- Model thinking 
- Computing for scientists

</details>


<details>
<summary> 
Week 2: August 31, Sep 4 – Environmental Data and Model Thinking  
</summary>

- Data scales 
- The row-format paradigm 
- Data recording, cleaning, and metadata 
- Model thinking: scientific and cultural contexts 

</details>


<details>
<summary>
Week 3: Sep 7, 11 – Data Exploration and Deterministic Functions  
</summary>

- Numerical and graphical data exploration 
- Functions 

</details>


<details>
<summary>
Week 4: Sep 14, 18 – The Dual-Model Paradigm, Uncertainty and Probability  
</summary>

- Essential probability theory basics 
- Uncertainty and stochastic processes 

</details>


<details>
<summary> Week 5: Sep 21, 25 – Frequentist Statistics </summary>

- Conceptual framework of the frequentist paradigm 
- Sample and population 
- Description and inference 

</details>

<details>
<summary> Week 6: Sep 28, Oct 2 – The Sampling Distribution and Confidence Intervals  </summary>

- Describing a sample vs. describing the sampling process 
- Frequentist interpretation of confidence intervals 

</details>

<details>
<summary> Week 7: Oct 5, 9 – Least Squares: Modeling Categorical and Continuous Data  </summary>

- Model fitting 
- Intro to t-tests, ANOVA, and simple linear models 

</details>


<details>
<summary> Week 8: Oct 12, 16 – Linear Models 1: Concepts  </summary>

- ANOVA
- simple and multiple linear regression. 

</details>


<details>
<summary> Week 9: Oct 19, 23 – Linear Models 2: Assumptions, Diagnostics, and Selection  </summary>

- Using models with real-life, messy data 

</details>




<details>
<summary>Week 10: Oct 26, 30 – Linear Models 3: Extending the Linear Model </summary>

- Survey of extended linear techniques: 
- generalized linear models
- mixed models
- generalized least squares
- other extensions of the linear model

</details>



<details>
<summary>Week 11: Nov 2, 6 – Frameworks: Frequentist, Bayesian, and Maximum Likelihood </summary>

- Contrasts among inferential paradigms

</details>

<details>
<summary> Week 12: Nov 9, 13 – Course Concepts Revisited</summary>

- Review of course concepts in the context of the course themes

</details>


<details>
<summary> Week 13: Nov 16, 20 – Where to go from here? </summary>

- Spatial Analysis
- Multivariate statistics
- Simulation
- Machine learning techniques

</details>
