---
title: "Essay with file upload"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Using the `dat_all` data frame, calculate the **total number of sampling sites** in which Gray Jays were observed.

Hint: What happens when you use the `sum()` function on a vector of Boolean values?

Q1 (1pt): Past the R code you used for the calculation into the text entry window.





Solution
========


```{r}
sum(dat_all$GRJA > 1)
```


Meta-information
================
extitle: Count of presences
extype: string
exsolution: nil
exname: exam name
exsection: section name
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 1