---
title: "One-tailed hypotheses"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---



```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



> Recall the Oregon birds dataset we've used in class.

Let's choose the Gray Jay and Brown Creeper as two example bird species.

Choose one of the terrain variables: elevation, slope, or aspect.

- NULL hypothesis: there is no difference in the habitat preferences of the Gray Jay and the Brown Creeper.

<div class="questionborder">

### Q1 2 pts.

- Propose a one-tailed alternative hypothesis about the habitat preferences of the two bird species.

NOTE: you don't need to look up habitat preference information about the two birds, I only care about how you phrase your hypothesis, not whether it is biologically/ecologically relevant!

</div>


 Polyscias fulva  Pseudospondias microcarpa
Solution
========

Acceptable answers can be in the form of:

Gray Jays occur [more, less] often in [low elevations, steep slopes, north-facing slopes] than Brown Creepers.






Meta-information
================
extitle: Two-tailed hypotheses
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 2

