---
title: "Confidence Intervals"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---



```{r, echo = FALSE, results = "hide"}
require(rmd.utils)
```



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


> Standard errors are used to calculate confidence intervals: the size of the interval is proportional to the value of the standard error.

Recall the formula for the standard error of the mean:

$SE_\bar{x} = \frac{SD_{sample}}{\sqrt{n}}$

Where $SD_{sample}$ is the sample standard deviation and $n$ is the sample size.



<div class="questionborder">

### Question

- (2 pts) What do you expect to happen to a confidence interval as the sample size increases?

</div>



Solution
========

- They shoud realize that confidence intervals shrink with larger sample sizes because of the sample size term in the denominator of the SE.





Meta-information
================
extitle: Confidence Intervals
extype: string
exsolution: nil
exname: week_05_reading_questions
exsection: week_05_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 2

