---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Distributions: Main Concepts"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(cowplot)
require(animation)
require(mfn.teaching.utils)
require(rmd.utils)
# source(here::here("rmd_tools", "find_file.R"))
# source(find_file("norm_density_plot.R"))

```




## Distribution Functions: Graphical Intuition

- Probability Density and Mass Functions
- Cumulative Density/Mass Functions
- Quantile Functions



## Probability Density/Mass Functions

Probability density is the y-value of the probability density curve for a given value of x.

- You can think of it as the height of a curve
- For *continuous* distributions, it is *not* equal to the probability of observing a particular value of x.







Demonstration of PDF and CDF using the Normal distribution.

Remember:

- *Density* = height of the curve at x.
- *Cumulative Density* = area under the curve, to the left of x



## PDF/CDF example

```{r}
norm_density_plot(-1.96)
```


## PDF/CDF example

```{r}
norm_density_plot(-1)
```


## PDF/CDF example

```{r}
norm_density_plot(-0.5)
```

## PDF/CDF example

```{r}
norm_density_plot(0)
```

## PDF/CDF example

```{r}
norm_density_plot(1)
```

## PDF/CDF example

```{r}
norm_density_plot(1.96)
```

## PDF/CDF example

```{r}
norm_density_plot(2.5)
```



- Later, we'll work on graphical intuition of probability functions.




## Recap of essentials:

### Distributions

1. They assign a *probability* to every *event* in a *sample space*.
1. We can use them as the *stochastic model* in the dual model paradigm.

### Probability essentials

1. Probabilities are non-negative
1. Probabilities of all events in sample space sum to 1.0
1. Independent events: joint probability is product of individual probabilities
