---
title: "Sampling and confidence intervals"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

> The width of confidence intervals is influenced by properties of both the population and the sampling process.



<div class="questionborder">

### Question

- (3 pts) Describe how the sampling process affects the width of confidence intervals.

</div>


Solution
========



Students should realize that larger sample sizes result in narrower confidence intervals.




Meta-information
================
extitle: Sampling and confidence intervals
extype: string
exsolution: nil
exname: week_06_reading_questions
exsection: week_06_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 3
