---
title: "Confidence interval interpretation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    # toc: TRUE
    # toc_float: TRUE
editor_options:
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

> Recall that we are *not* 95% certain that a 95% confidence interval we calculate contains the true value.


<div class="questionborder">

### Question

- (8 pts) Explain what a frequentist 95% confidence interval really means.  Make sure you cover any relevant assumptions of the Frequentist paradigm.  You answer must be in non-technical language.  Imagine you were explaining confidence intervals to an audience of teenagers.

HINT: Your explanation is likely to be more successful if you use an example experiment, or describe your answer in the context of a real-life scenario rather than a purely theoretical explanation.

</div>


Solution
========



Students should realize that larger sample sizes result in narrower confidence intervals.




Meta-information
================
extitle: Confidence Interval Interpretation
extype: string
exsolution: nil
exname: week_06_reading_questions
exsection: week_06_reading_questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15
expoints: 8
