---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Loading an RProject"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(list.files(path = here::here("formatting"), pattern = "rmd_functions.R", recursive = TRUE, full.names = TRUE))

```

# Background

You've created a great directory structure for your coding project, and you had RStudio create an RProject file to keep it all organized!

But to take advantage of the benefits of using the RProejct paradigm (like using package `here`) you have to tell RSTudio to load the project!




# Load an RProject when you open RStudio

The best practice is to load your RProject as soon as you open up RStudio.

The easiest way to accomplish this is to open RStudio by double-clicking on the `*.RProj` file in your RProject directory:


```{r, echo = FALSE, results='asis'}
expandable_html_image("open_rproject_from_file_browser.PNG")
```






