---
title: "Sample Space and Events 2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


> A sample space is the ....

Suppose it is a beautiful fall day and you are sitting underneath three oak trees: Bur oak (*Quercus. macrocarpa*), Northern Red Oak (*Q. rubra*), and white oak (*Q. alba*).

- There are approximately the same number of acorns from each species.

You collect an acorn, walk a short distance and collect a second acorn.


<div class="questionborder">
### Q1 2 pts 

- What is the probability that *both* acorns are *Q. alba*?

</div>


Solution
========


A = alba

M = macrocarpa

R = rubra

- The student should realize that the order *might* be important since the acorns were collected serially, but that there is only one way to collect two A's.

The sample space is:

```{r, echo = FALSE}
data.frame(
  first_acorn = rep(LETTERS[1:3], each = 3),
  second_acorn = rep(LETTERS[1:3], 3))
```


- There are 9 possible collections:
  - The sample space has 9 elements.

- There is only one way to choose two acorns that are the same species.
- The probability is approx 1 / 9 or `r round(1/9, 2)`



Meta-information
================
extitle: Sample space and events 2
extype: string
exsolution: nil
exname: week 4 reading questions
exsection: week 4 reading questions
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 2
