---
title: "[ECo 602 - Analysis of Environmental Data](/eco_602_634_2020/index.html){target='_blank'}"
subtitle: "Week 12: In-class Final Project Data"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  # epuRate::epurate:
  #   toc: TRUE
  #   number_sections: FALSE
  # code_folding: "hide"
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
require(mfn.teaching.utils)
require(rmd.utils)
```

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

```{r build_assignment_page, eval=FALSE, include=FALSE}
require(rmd.utils)
```



# Learning Objectives

- Import data for the final project.
- Preliminary graphical exploration.



# Data Files

You'll need to download the `delomys.csv` data file from the Final Project Data section of the course GitHub page https://michaelfrancenelson.github.io/eco_602_634_2020/.

- Make sure you save the file in the "data" subdirectory of your main course RProject directory.
- Load the data into a data.frame in R.  You should call your data frame `dat_delomys`.



# Numerical and Graphical Exploration

## Take a peek at the data.
 
- Which data columns hold continuous data?
- Which data columns hold categorical data?


## Graphical exploration

- Plot a histogram of body mass.
- Plot a histogram of body length.
- Plot a scatterplot of body length (on the y-axis) and body mass (on the x-axis)


# Questions

As a group, submit a document with answers to the following:

1. Which data columns hold continuous data?
2. Which data columns hold categorical data?

Include the following figures in your document:

1. Scatterplot of body length and body mass.
2. Conditional boxplot of body mass grouped by sex.
3. Conditional boxplot of body mass grouped by species.
4. Conditional boxplot of body mass grouped by species and sex.



```{r data_analysis, echo=FALSE, eval=FALSE}

here::here("data", "delomys.csv")

dat_delomys = read.csv(here::here("data", "delomys.csv"))
head(dat_delomys)


plot(body_length ~ body_mass, data = dat_delomys)
boxplot(body_mass ~ binomial, data = dat_delomys)
boxplot(body_mass ~ sex, data = dat_delomys)
boxplot(body_mass ~ sex * binomial, data = dat_delomys)



```

