---
title: "Standard error of the mean function"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}

cat(readLines(here::here("formatting", "css", "styles.css")))
require(palmerpenguins, quietly = TRUE)
sse_mean = function(x) 
{
  # You have to use na.rm = TRUE
  sample_sd = sd(x, na.rm = TRUE)
  
  # There are many ways to calculate the number of non-missing
  # elements in x.
  n = sum(!is.na(x))
  
  return (sample_sd / sqrt(n))
}
```



<div class="instructionsborder">

#### Instructions

> Recall your `sse_mean()` function. Your function should accept a single numeric vector and return a numeric value:


```{r eval = FALSE, results='asis'}
# Function template
sse_mean = function(x)
{
  
  # ... your code here ...
  
  return(sse)  
}
```

You can compare your results to these to make sure that your standard error
function is working correctly:

```{r, results='hold'}
sse_mean(penguins$body_mass_g)
sse_mean(mtcars$mpg)
```

</div>


<div class="questionborder">

### Q1 3 pts.

- Paste the code you used to define your `sse_mean()` function into the text input.

- include the following line before your function definition:

```{r eval = FALSE}
rm(list = ls())
```

and the following two test lines after your function:

```{r eval=FALSE}
sse_mean(penguins$body_mass_g)
sse_mean(mtcars$mpg)
```


</div>



Solution
========

Their standard error function should return:

- 43.36 for the penguin body mass.
- 1.065 for the car mpg.




```{r, results='hold'}
sse_mean(penguins$body_mass_g)
sse_mean(mtcars$mpg)
```

My implementation:




```{r reuslts}
require(palmerpenguins)

sse_mean = function(x) 
{
  # You have to use na.rm = TRUE
  sample_sd = sd(x, na.rm = TRUE)
  
  # There are many ways to calculate the number of non-missing
  # elements in x.
  n = sum(!is.na(x))
  
  return (sample_sd / sqrt(n))
}

sse_mean(penguins$body_mass_g)
sse_mean(penguins$flipper_length_mm)

```


### Grading guide

Full credit if

- They had to include the `na.rm = TRUE` argument to `sd()` in their function definition.
  - The function won't run on the penguin data if they didn't include `na.rm = TRUE`.
- Their function must accurately adjust the number of samples when there are missing data.
  - They can implement this in many ways, some more complicated than others.  As long as their method works they shoudl get full credit.  They had a similar problem in an earlier lab.

Deductions:

- Deduct 1 point if they didn't include `na.rm = TRUE`.
- Deduct 1 point if they didn't adjust n for NA values.


Some possible ways to adjust n include:

```{r eval = FALSE}
n = length(na.omit(x))
n = sum(!is.na(x))
n = length(x)-length(x[is.na(x)])

```



Meta-information
================
extitle: Standard error of the mean function

extype: string

exsolution: nil

exname: lab_06

exsection: lab_06

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 15

expoints: 3



