---
title: "Two-sample resampling function"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


```{r, echo = FALSE, results = "hide"} 
  
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructionsborder">

#### Instructions

> Your `two_group_resample()` function should accept three arguments:

1. `x`: A numeric vector
1. `n_1`: The number of values (with replacement) to take from `x` for the first group.
1. `n_2`: The number of values (with replacement) to take from `x` for the second group.


Your function should return:

- A single numeric value: the difference in means between the two groups

```{r echo = TRUE, eval = FALSE, results='asis'}

two_group_resample = function(x, n_1, n_2) 
{
  
  # ... your code here ...
  
  return(difference_in_means)
}

```
</div>





<div class="questionborder">

### Q1 4 pts

Paste the code you used to define your `two_group_resample()` function into the text input.

</div>



Solution
========

Some possible implementations:

```{r}

two_group_resample = function(x, n_1, n_2) 
{
  dat_1 = sample(x, n_1, replace = TRUE)
  dat_2 = sample(x, n_2, replace = TRUE)
  return(mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE))
}

two_group_resample = function(x, n_1, n_2) 
{
  m_1 = mean(sample(x, n_1, replace = TRUE), na.rm = TRUE)
  m_2 = mean(sample(x, n_2, replace = TRUE), na.rm = TRUE)
  return(m_1 - m_2)
}


```






### Grading Guide

A common mistake students made was to keep code that overwrites the values of the inputs:

```{r common_wrong_answer, eval = FALSE}
two_group_resample = function(x, n_1, n_2)
{
  x = dat_pen$flipper_length_mm
  n_1 = 52
  n_2 = 108
# ...
} 
```

- They should receive 2 out of 4 points if the rest of the code is correct.  I've used the following response:


<div = class="hintborder">
Your code is almost correct.
<br>

Notice that within the function body, you overwrite the values of x, n_1, and n_2:

<ul>
<li> x = dat_pen$flipper_length_mm </li>
<li>  n_1 = 52 </li>
<li>  n_2 = 108 </li>
</ul>

This means that no matter what arguments you try to use when you call the function, your function will always overwrite them to these values.
</div>


Grading test case.  Should return approximately -0.56 if students called `sample()` in approximately the same way as the reference.

- Returning a different value doesn't automatically mean their implementation is wrong!


```{r}
require(palmerpenguins)
dat_pen = droplevels(subset(penguins, species != "Gentoo"))

two_group_resample = function(x, n_1, n_2)
{
  dat_1 = sample(x, n_1, replace = TRUE)
  dat_2 = sample(x, n_2, replace = TRUE)
  return(mean(dat_1, na.rm = TRUE)-mean(dat_2, na.rm = TRUE))
}


 


require(palmerpenguins)
dat_pen = droplevels(subset(penguins, species != "Gentoo"))
set.seed(12345)
two_group_resample(dat_pen$flipper_length_mm, 68, 152)
```


Some possible w




Meta-information
================
extitle: Two-sample resampling function

extype: string

exsolution: nil
exname: lab_06

exsection: lab_06

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12

expoints: 4


