---
title: "Resampling penguin data: histogram"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 

```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructionsborder">

#### Instructions

Use your `two_group_resample()` function along with a loop to generate 2000 resampled
differences of means for flipper length between the two penguin species.

- Hint: see the **Resampling simulations** section in the lab
walkthrough for an example of how to do this.

</div>

<div class="questionborder">

### Q1 2 pts

Upload a histogram of the resampled differences of means. 

</div>



Solution
========


```{r}
require(palmerpenguins)
dat_pen = droplevels(subset(penguins, species != "Gentoo"))

two_group_resample = function(x, n_1, n_2) 
{
  dat_1 = sample(x, n_1, replace = TRUE)
  dat_2 = sample(x, n_2, replace = TRUE)
  
  diff_simulated = 
    mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
  
  return(diff_simulated)
}


n = 2000
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample(dat_pen$flipper_length_mm, 68, 152)
  )
}
hist(mean_differences)
sum(abs(mean_differences) > 5.8)
```



Meta-information
================
extitle: Resampling penguin data: histogram

extype: string

exsolution: nil

exname: lab_06

exsection: lab_06

exextra[essay,logical]: TRUE

exextra[essay_format,character]: noinline

exextra[essay_required,logical]: FALSE

exextra[essay_fieldlines,numeric]: 5

exextra[essay_attachments,numeric]: 1

exextra[essay_attachmentsrequired,logical]: TRUE

expoints: 2


