---
title: "Resampling: p-value"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(palmerpenguins)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructionsborder">

#### Instructions

Recall this interpretation of a Frequentist p-value:

> How often would I expect to see a result as extreme, or more extreme, if the null hypothesis were true?

The p-value from the t-test for the difference in mean flipper length was very, very small:

- 6.049e-08

That's less than once in 10 million!

</div>


<div class="questionborder">

### Q1: 2 pts

- Given a p value of less than 1 per 10 million, how many simulations
do you think you would have to do to see a difference in
mean flipper length equal to or greater than 5.8 mm?

</div>

<div class="hintborder">

- Hint: You don't need to run any simulations to answer this question.

- Hint: The number may be very large, I don't recommend trying to run more 
than 10 thousand simulations unless you want to wait a very long time for results!

</div>


Solution
========

Students should realize that they would probably need to carry out
millions of simulations before they saw an *interesting* result!



Meta-information
================
extitle: Resampling: p-value

expoints: 2

extype: string

exsolution: nil

exname: lab_06

exsection: lab_06

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 5

