---
title: "Resampling"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
require(palmerpenguins)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructionsborder">

#### Instructions


<div class="deemphasize">
- Choose another variable of penguin data from `dat_pen`.
- Create a boxplot of the data from your chosen column in `dat_pen`, grouped by species. 
  - See the walkthrough for info.
</div>

<div class="emphasis">
- Use `aggregate()` to calculate the group means and the difference in means.
  - Write the difference in means into a variable called `diff_crit`
- Conduct a t-test and observe the p-value.
</div>

<div class="deemphasize">
- Conduct a resampling test with 1000 repetitions using your `two_group_resample()` function.
</div>

</div>

<div class="questionborder">

### Q1: 4 pts

- Interpret the p-value from the t-test output in plain, non-technical English that
a non-statistician would understand.


</div>



Solution
========

Answers will vary...



Meta-information
================

extitle: resampling

expoints: 4

extype: string

exsolution: nil

exname: lab_06

exsection: lab_06

exextra[essay,logical]: TRUE

exextra[essay_format,character]: editor

exextra[essay_required,logical]: TRUE

exextra[essay_fieldlines,numeric]: 12



