---
title: "Logical subsetting - intermediate q2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
require(rmd.utils)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Run the following code to create a large vector containing randomly generated integers between 1 and 12:

<br>
```{r eval=FALSE}
n = 12345
vec_1 = sample(12, n, replace = TRUE)
head(vec_1)
```
<br><br>


Use the function `length()` to determine how many elements are in `vec_1`.
 
<br>

Now, run the following line to check how many entries have the value `3`:

```{r eval=FALSE}
sum(vec_1 == 3)
```

<br><br>

Finally, run the following code several times taking note of how many `3` entries appear each time you run it.

<br><br>

```{r, eval=FALSE}
n = 10
vec_1 = sample(12, n, replace = TRUE)
paste0("Sum of elements with value 3: ", sum(vec_1 == 3))
```
<!-- print(vec_1) -->

<br><br>

Did you get different counts of `3` entries each time?


<div class="questionborder">

### Q1 3 pts

- Considering the different vectors generated each time, explain why using a logical test is a safe way to select entries with a value of `3`. 

### Q2 5 pts

-Explain why manually choosing the indices by visual inspection, is problematic.  You may want consider re-usability of code, working with different sized data sets, and sharing code with collaborators.

</div>





Solution
========

Q1: The values of 3 won't be in the same position, using a logical test will always return the correct position indices.

Q2: Many possible answers:

- Vectors may be too long to visually inspect
- Visual inspection is error-prone
- Hard-coding indices is generally bad practice
- lots of other possibilities




Meta-information
================
extitle: logical subsetting intermediate 1
extype: string
exsolution: nil
exname: lab 02
exsection: lab 02
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 8

