---
title: "basic loop question 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```



You may want to review the *for-loop* example in the lab walkthrough.


```{r, eval=FALSE}
for (i in 1:10)
{
  print(i)
}
```

<br>

Modify the code in the body of the loop to print out a message like "This is loop iteration: 1" for each run through the loop. 

<br>

- Hint:  use the `print()` and [`paste()` or `paste0()`] functions.
- Hint: review the *nesting functions* example in the lab walkthrough.

<br>

<div class="questionborder">
### Q1 (3 pts)

- Paste your loop code into the text input field.

</div>

Solution
========

Students should produce something like:
```{r, echo = TRUE}
for (i in 1:10)
{
  print(paste0("This is iteration ", i, "."))
}
```

If they want to be fancy they can use `sprintf()`:
```{r}
for (i in 1:10)
{
  print(sprintf("This is iteration %s.", i))
}
```




Meta-information
================
extitle: basic loop question 1
extype: string
exsolution: nil
exname: lab 02
exsection: lab 02
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 3

