---
title: "intermediate loop question 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---



Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

You may want to review the *for-loop* example in the lab walkthrough.


```{r, eval=FALSE}
for (i in 1:10)
{
  print(i)
}
```

Note that the loop runs through exactly 10 iterations...

What if you wanted the loop to execute an arbitrary number of times?

<br>

- Create a variable, `n`, that contains an integer value.
- Modify the code for the loop so that it runs `n` times.

<br>

<div class="questionborder">
### Q1 (2 pts): 

- Paste the code for the modified loop that executes `n` times into the text window below.

</div>


Solution
========

Students should produce something like:
```{r, echo = TRUE}
n = 17
for (i in 1:n)
{
  print(paste0("This is iteration ", i, "."))
}
```





Meta-information
================
extitle: intermediate loop question 1
extype: string
exsolution: nil
exname: lab 02
exsection: lab 02
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 2

