---
title: "intermediate loop question 2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 

```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

### Generate a random vector and print the contents with messages:

<br>
 
- Create an integer variable, `n`, that holds the value `17`.
- Write code to create a vector called `vec_1` of length `n`.  `vec_1` should contain [pseudo]randomly generated integers between 1 and 10.

<div class="hintborder">
- Hint:  Check out the help entry and consult Dr. Google on how to use the `sample()` R function.
- Hint:  Take a look at the code I provided in an earlier question for an example of how to create a vector of random values.
</div>

Now use the code you used to create a loop the executes `n` times to print out each element of `vec_1`.  Each iteration through the loop should print the corresponding element of `vec_1`.

<br>

Your output should look something like this:

<div class="outputborder">
```{r, echo=FALSE}
n = 6
vec_1 = sample(10, n, replace = TRUE)
for (i in 1:n)
{
  cat(sprintf(
    "The element of %1$s at index %2$s is %3$s.\n",
    "vec_1", i, vec_1[i]))
}
```
</div>
<div class="hintborder">
Hint: Think of what code you'll need to include within the body of the loop to display:
1. index number
1. The value of `vec_1` at the index.
</div>


<br>

<div class="questionborder">

### Q1 4 pts

- Paste the code you used to create the `n`, `vec_1`, and the loop in to the text entry box.
</div>


Solution
========

Students should produce something like:
```{r, echo = TRUE}
n = 6
vec_1 = sample(10, n, replace = TRUE)


# paste() version
for (i in 1:n)
{
  print(paste0(
    "The element of vec_1 at index ",
    i, 
    " is ",
    vec_1[i]))
}

# sprintf version
for (i in 1:n)
{
  cat(sprintf(
    "The element of %1$s at index %2$s is %3$s.\n",
    "vec_1", i, vec_1[i]))
}


```





Meta-information
================
extitle: intermediate loop question 2
extype: string
exsolution: nil
exname: lab 02
exsection: lab 02
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 4

