---
title: "custom function 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


## Objective

<div class="redborder">
Write a function `create_and_print_vec()`.
</div>

- You've created a loop that can print a message with the value of each element of a vector of arbitrary length.
- Now you'll wrap it all into a custom function.
- Review the lab materials about writing custom functions in the lab walkthrough.



### Function Arguments

<div class="blackborder">
Your function should take three integer arguments, `n`, `min`, and `max`.

- `n` has no default value.
- `min` has a default value of 1.
- `max` has a default value of 10.
</div>

### Function Output

<div class="blackborder">
Your function needs to do the following:

- Create a vector of `n` random integers between the values of `min` and `max`.
- Loop through the elements of the vector and print a message with the index of the element and its value.
</div>

<div class="outputborder">
Here's a skeleton:

```{r eval=FALSE}
create_and_print_vec = function(n, min = , max =)
{
  # Function body goes here
}
```
</div>

<div class="hintborder">
- Hint: You've already written code that accomplishes *almost* all of these tasks.
- Hint: You've used a single number as the first argument to `sample()`.  Look up the R help entry to see what other kinds of values the `x` argument of `sample()` can accept.
</div>

<br>

Your function should create output like this with default values for `min` and `max`:

<div class="outputborder">
```{r, echo=FALSE}
create_and_print_vec = function(n, min = 1, max = 10)
{
  vec_1 = sample(min:max, n, replace = TRUE)
  for (i in 1:n)
  {
    print(paste0(
      "The element at index ",
      i, 
      " is ",
      vec_1[i]))
  }
}
create_and_print_vec(10)
```
</div>
<br>

When you use `min = 100` and `max = 2000` your output should resemble:

<div class="outputborder">
```{r}
create_and_print_vec(10, min = 100, max = 2000)
```
</div>
<br>

<div class="questionborder">

### Q1 10 pts

- Paste the code you used to build your function.
  - To receive full credit your code must run without error on a new R session and produce output similar to the examples given in the instructions.

</div>

Solution
========

Students should produce something like:

```{r, echo = TRUE}

create_and_print_vec = function(n, min = 1, max = 10)
{
  vec_1 = sample(min:max, n, replace = TRUE)
  for (i in 1:n)
  {
    print(paste0(
      "The element at index ",
      i, 
      " is ",
      vec_1[i]))
  }
}

create_and_print_vec(10)
create_and_print_vec(10, min = 100, max = 2000)
```





Meta-information
================
extitle: custom function 1
extype: string
exsolution: nil
exname: lab 02
exsection: lab 02
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expoints: 10

