---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Detailed Course Schedule"
author: "Michael France Nelson"
date: "Fall 2020"
output:
  pdf_document:
    toc: true
    number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    toc_float: TRUE  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

tue_start = as.Date("2020-08-25")
thu_start = tue_start + 2
dat_fmt = "%b %d"

dat_str = function()
return(paste0(format(tue_start, dat_fmt), ", ",
       format(thu_start, dat_fmt)))

dat_str()

```




## Week 1: `r dat_str()` {.tabset}
```{r increment_date, include = FALSE}
tue_start = tue_start + 7; thu_start = tue_start + 2
```
<a name="week01"></a>
**Introductions, Why Model? **


### Key Concepts

- Introduction to course 
- Course themes 
- Model thinking 
- Computing for scientists 


### Readings 

- Epstein: Why Model? 
- McGarigal ch. 1 
- Bolker Chapter 1: sections 1.1, 1.2, 1.3 


### Lectures 

- Navigating Course Moodle 
- Model Thinking 1: What is a Model? [Note: this will be covered in the first course meeting on Tuesday Sep. 25]
- Computing for Scientists 
- R and RStudio 


### Individual Assignments

- Learning objectives assessment: Due Friday Aug 28 by midnight
  - Grade is based on completion, not correct answers.
- Software setup: Due Monday Aug 31st
- DataCamp: Intro to R - Due Friday Sep. 4th


### Zoom Meetings

#### Tuesday: Course introduction, structure, and overview 

- Course structure and format 
- Course themes 
- My expectations 
- Learning objectives
- What is a Model?


#### Thursday: Discussion 

- Your expectations  
- Model Thinking and Course Themes
- Model thinking in-class activity


 


##  Week 2: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week02"></a>
**Environmental Data and Model Thinking**

### Key Concepts

- Data scales 
- The row-format paradigm 
- Data recording, cleaning, and metadata 
- Model thinking: scientific and cultural contexts 


### Lectures

- Data Scales
- Model Thinking: population and sample
- Model Thinking: a frequentist perspective


### Readings 

- Bang, Megan, Ananda, Marin, and Medin
- McGarigal Chapter 2


### Individual Assignments

- Course GitHub Repositlry Setup


### Zoom Meetings

#### Tuesday

#### Thursday



## Week 3: `r dat_str()`{.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week03"></a>
**Data Exploration and Deterministic Functions **

### Key Concepts

- Numerical data exploration
- Graphical data exploration 
- Functions 


### Readings 

- Zuur Chapter 4.1.  Focus on the sections below.  We’ll cover some of the other plot types later. 
  - Boxplots and conditional boxplots 
  - Histograms 
  - Scatterplots 
  - Pair plots 
- McGarigal Chapter 3: Data Exploration, Screening & Adjustments
  - sections 1 – 5.  This covers much of the same material as the Zuur reading, but the writing style is perhaps more approachable.
- McGarigal Chapter 4: Deterministic Functions
  - Sections 1-3, 5
  - section 4: 
    - Glance through this section.  Try not to get caught up in the mathematical details.
    - Focus on the figures and the shapes of the functions.
    - We’ll cover the details of selected functions in lecture and discussion. 


### Lectures 

- Functions: function types and graphical intuition 
- Functions: understanding formulae
- Describing data with deterministic functions


### Individual Assignments

- Deterministic Functions

### Zoom Meetings

#### Tuesday

#### Thursday








## Week 4: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```
<a name="week04"></a>
**The Dual-Model Paradigm, Uncertainty and Probability**

### Key Concepts

- Essentials of probability theory
- Uncertainty
- Stochastic processes 



### Readings

- Bolker ch. 1 section 1.6
- McGarigal 5a



### Lectures 

- Probably distributions 1: general concepts
- Discrete distributions
- Continuous distributions



### Individual Assignments 

- Dual Models


### Zoom Meetings

#### Tuesday

#### Thursday



## Week 5: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week05"></a>
**The Frequentist Paradigm**


### Key Concepts

- Sample and population
- Statistics and parameters
- Repeated sampling



### Readings 

Bolker ch 1, section 1.4
McGarigal Ch 5b, 5c, 5d
Jorge Luis Borges: The Library of Babel



### Lectures 

- The frequentist perspective
- Stochastic processes and repeated sampling



### Individual Assignments 

- Continue Dual Models assignment.




### Zoom Meetings

#### Tuesday

#### Thursday




## Week 6: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week06"></a>
**The Sampling Distribution and Confidence Intervals**


### Key Concepts

- Describing a sample vs. describing the sampling process 
- Frequentist interpretation of confidence intervals 


### Readings 

- McGarigal Ch 5d



### Lectures 

- Sample statistics and the sampling distribution
- The Frequentist confidence interval: definitions, uses, and abuses


### Individual Assignments 

- Frequentist Concepts



### Zoom Meetings

#### Tuesday

#### Thursday



## Week 7: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week07"></a>
**Least Squares: Modeling Categorical and Continuous Data**


### Key Concepts

- Model fitting 
- Null and alternative hypotheses
- Overview of simple inferential models
  - T-tests
  - ANOVA
  - Simple linear models 


### Readings 

McGarigal Ch 6
McGarigal Ch. 7
Lecture Videos
How do we measure 'error'?
Optimizing

### Lectures 

### Individual Assignments 

- Using Models I


### Zoom Meetings

#### Tuesday

#### Thursday




## Week 8: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week08"></a>
**Linear Models 1: Key Concepts**



### Key Concepts

- Modeling categorical data: ANOVA
- Modeling continuous data: linear regression




### Readings 

- Bolker Ch. 9, sections 9.1 - 9.2
- Zuur Ch. 5.1 - 5.2




### Lectures 

- Functions: continuity and local linearity
- Data Transformations
- 



### Individual Assignments 

- continue Using Models I assignment


### Zoom Meetings

#### Tuesday

#### Thursday




## Week 9: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week09"></a>
**Linear Models 2: Assumptions, Diagnostics, and Selection **


### Key Concepts

- Fitting models to real, messy data
- Testing assumptions



### Readings 

- Zuur Ch. 5.1 - 5.2



### Lectures 

- When assumptions aren't met
- When assumptions aren't reasonable



### Individual Assignments 

- Using Models II



### Zoom Meetings

#### Tuesday

#### Thursday





## Week 10: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week10"></a>
**Linear Models 3: Extending the Linear Model **

### Key Concepts

- Survey of extended linear techniques: 
  - generalized linear models 
  - mixed models 
  - generalized least squares 
  - other extensions of the general linear model 


### Readings 

- Bolker 9.3, 9.4, 9.5
- Zuur 6, 7.1 - 7.4



### Lectures 

- Extending the general linear model 1: discrete data
- Extending the general linear model 2: abnormal errors
- Extending the general linear model 3: nonlinearity



### Individual Assignments 

- continue Using Models II assignment

### Zoom Meetings

#### Tuesday

#### Thursday



## Week 11: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week11"></a>
**Frameworks: Frequentist, Bayesian, and Maximum Likelihood **

### Key Concepts

- Contrasts between the frameworks



### Readings 

- McGarigal Ch 6 - 8



### Lectures 

- Bayesian and Frequentist paradigms
- Likelihood: concepts and intuition
- Analytical solutions, numerical methods, and simulation



### Individual Assignments 

- Using Models III


### Zoom Meetings

#### Tuesday

#### Thursday






## Week 12: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week12"></a>
**Course Key Concepts Revisited**

### Key Concepts

### Readings 

### Lectures 

### Individual Assignments 

- continue Using Models III assignment

### Zoom Meetings

#### Tuesday

#### Thursday




## Week 13: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
<a name="week13"></a>
**Where to go from here?**

### Key Concepts

- Spatial Analysis 
- Multivariate statistics 
- Simulation 
- Machine learning techniques 



### Readings 

### Lectures 

### Individual Assignments 

- continue Using Models III assignment


### Zoom Meetings

#### Tuesday

#### Thursday
