---
title: "ECo 602 - Analysis of Environmental Data"
author: "Michael France Nelson"
subtitle: "Mini-Lecture Schedule"
date: "Fall 2020"
output:
  # pdf_document:
  #   toc: true
  #   number_sections: TRUE
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "eco_602_2020.css")
    toc: TRUE
    toc_float: TRUE  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

tue_start = as.Date("2020-08-25")
thu_start = tue_start + 2
dat_fmt = "%b %d"

dat_str = function()
return(paste0(format(tue_start, dat_fmt), ", ",
       format(thu_start, dat_fmt)))

dat_str()

```


#  Week 1 – Introductions, Why Model?: `r dat_str()` {.tabset}
```{r increment_date, include = FALSE}
tue_start = tue_start + 7; thu_start = tue_start + 2
```
- Navigating Course Moodle
  - Where to find syllabus
  - GitHub Pages
  - Weekly section organization
    - Mini-lectures and lecture notes
    - Assignments
      - Moodle Quiz questions
        
        
- Model Thinking: What is a Model?
  - Dog screen door video
  - Multiple models
  
  
- Computing for Scientists
  - Common issues: where is my file?
  - GUI and text-based interfaces
  - Paths
    - more of this in reading and individual assignment


- R and RStudio
  - Why R (or scripting) instead of menu-driven analyses
    - Reproducibility
  - Demo of simple tasks
  - Demo of an analysis
  - Demo of simple and complex graphics




#  Week 2 - Environmental Data and Model Thinking: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```
- Model Thinking: Communicating Science
  - McGarigal's Bird examples

- Data Scales
  - Examples for each
    - continuous and discrete
    - ratio and interval
    - ordinal and categorical
    - converting among data scales

- Data Conventions
  - Recording and quality control
  - Metadata
  - Row-data paradigm




# Week 3 - Data Exploration and Deterministic Functions: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```

- Why Data Exploration?
  - Numerical exploration
    - Why should we do it?
    - Statistics and Parameters
    - Measures of center and spread
  - Graphical exploration
    - Why should we do it?
    - What aspect(s) of the data do we want to explore?>

- Important Plot Types
  - What do we want to learn?
  - Plots that display summaries
  - Plots that show raw data
  
- Functions
  - What is a function?
  - Formulas and equations
  - Important classes of functions
    - Linear
    - Polynomial and rational
    - Exponential and logarithmic
    - Periodic
    - Combination
  
  
- Deterministic Functions
  - Always start with a line!
  - Important types for environmental data:
    - Give a mechanistic and/or phenomenological example of:
      - linear
      - periodic
      - Combination functions, e.g. Ricker-type





# Week 4: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE, echo =FALSE}
```




## Live Sessions


### Tuesday

### Thursday



## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 5: Sep 22, 24  {.tabset} -->
# Week 5: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```


## Live Sessions


### Tuesday

### Thursday


## Recorded Mini Lectures 

## Readings 

## Assignments 



<!-- # Week 6: Sep 29, Oct 1  {.tabset} -->
# Week 6: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```

## Live Sessions



### Tuesday

### Thursday


## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 7: Oct 6, 8  {.tabset} -->
# Week 7: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```

## Live Sessions



### Tuesday

### Thursday



## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 8: Oct 13, 15  {.tabset} -->
# Week 8: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```

## Live Sessions



### Tuesday

### Thursday


## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 9: Oct 20, 22  {.tabset} -->
# Week 9: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```


## Live Sessions



### Tuesday

### Thursday



## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 10: Oct 27, 29  {.tabset} -->
# Week 10: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```


## Live Sessions



### Tuesday

### Thursday


## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 11: Nov 3, 5  {.tabset} -->
# Week 11: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```


## Live Sessions


### Tuesday

### Thursday




## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 12: Nov 10, 12  {.tabset} -->
# Week 12: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```


## Live Sessions



### Tuesday

### Thursday


## Recorded Mini Lectures 

## Readings 

## Assignments 




<!-- # Week 13: Nov 17, 19  {.tabset} -->
# Week 13: `r dat_str()` {.tabset}
```{r ref.label = "increment_date", include = FALSE}
```


## Live Sessions



### Tuesday

### Thursday



## Recorded Mini Lectures 

## Readings 

## Assignments 



