---
title: "Histograms of sampling site terrain"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Instructions:

1. Create histograms for the three terrain variables: elevation, slope, and aspect.
1. Create a single plot including all three histograms and save it as a .png image file.


<div class="hintborder">

- Hint: you can use `par(mfrow = c(3, 1))` to create a figure with three panels arranged in a single column.
- Hint: `par(mfrow = c(1, 3))` will create a figure with three panels arranged in a single row.
- Hint: Choose dimensions for your output file so that the individual histograms have an appropriate aspect ratio.
- Hint: You may notice some pacularities with the rightmost bin in the aspect histogram.  Consider the units in which aspect is measured and check out the `breaks` argument for `hist()`.

</div>


<div class="questionborder">

### Q1 4 pts 

- Upload an image file of your figure.

</div>



Solution
========

3 pts if students create all the histograms, but don't fix the rightmost bin problem in the aspect plot.

The aspect histogram issue:

- It's due to an artifact of the data format and R-s binning algorithm
- Degrees are in the interval 0 to 360
- The last bin contains the interval 350 - 400
- `bins

```{r aspect_hist, fig.asp=1/3}
dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

{
  par(mfrow = c(1, 3))
  hist(
    dat_all$elev, 
    xlab = "elevation (m)",
    main = "Sample Site Elevation")
  hist(
    dat_all$slope,
    xlab = "slope (percent)",
    main = "Sample Site Slope")
  hist(
    dat_all$aspect, 
    xlab = "aspect (degrees)",
    main = "Sample Site Aspect",
    breaks = seq(0, 360, length.out = 12))
}
```







Meta-information
================
extitle: Histogram of aspect
extype: string
exsolution: nil
exname: data exploration det. funs
exsection: data exploration det. funs
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
expoints: 6
