---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Hypothesis Testing 2"
author: "Michael France Nelson"
date: "Fall 2020"
output:a
    beamer_presentation:
    pandoc_args: !expr paste0(here::here("formatting", "beamer", "eco_602_2020_beamer.yaml"))
    highlight: tango
    # theme: "default"
    colortheme: "spruce"
    fonttheme: "serif"
    slide_level: 2
    incremental: false
classoption: t
header-includes:
  \input{`r here::here("formatting", "beamer", "eco_602_2020_headers_tikz.tex")`}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Key Concepts and Terms

- Hypotheses: null and alternative




## Hypotheses

### Hypothesis testing is central to Frequentist statistics.

What is hypothesis testing?

What is a hypothesis?




## Null Hypotheses




## Alternative Hypotheses










## Bolker's seeds example

Bolker's writing is ... difficult.  Let's decipher the seed predation example.

### The question (as originally posed in the reading): Are there different predation rates on different plants' seeds?

Two example species (the largest and smallest from a set of 8) used in the example:
- Polyscias fulva (pol: seed mass < 0.01 g)
- Pseudospondias microcarpa (psd: seed mass ≈ 50 g).

### The null hypothesis: seed predation rates are equal among species



## Bolker's seeds example: exploration of H_0

### Why is this a good null hypothesis?

- It is intuitive: if we find good *evidence* that the predation rates are different, it's reasonable to conclude there is a  *true difference*.
- It's simple
- It's easy to test!



## Bolker's seeds example: exploration of H_0

### Where are the opportunities for *improvement*?

- Is it reasonable to expect that predation rates should ever be equal?
- If we find good evidence that rates are different, what does that tell us?




## Bolker's seeds example: exploration of H_0

### If the null hypotheses were true, what would we expect to observe?

### What differences in rates could we expect to observe by chance alone?



## Bolker's seeds example: exploration of H_0

### Simulations in R

Let's do some simulations to see what we might expect to observe *by chance alone*.

For the simlations, let's use the following parameters

- 200 observation stations
- Predation rates for the two species are equal: 50%
- For each station there is a binary outcome: taken or not taken

### This sounds a lot like a binomial distribution problem!




```{r}
n_simulations = 1000
n_stations = 200
pred_rate = 0.5


set.seed(12345)
sim_taken = 
  data.frame(
    pol = rbinom(n = n_simulations, size = n_stations, prob = pred_rate),
    psd = rbinom(n = n_simulations, size = n_stations, prob = pred_rate))
sim_not_taken = n_stations - sim_taken




sim_taken_rate = sim_taken / sim_not_taken
hist(apply(sim_taken_rate, 1, FUN = function(x) x[1] / x[2]))


head(sim_data)
```



```{r}

sim_i = 4
ft_i = fisher.test(
  rbind(
  sim_data[sim_i, ],
  n_stations - sim_data[sim_i, ]))

ft_i$p.value
ft_i$conf.int
ft_i$estimate


fisher.test(cbind(c(109, 91), c(105, 95)))
fisher.test(rbind(c(109, 91), c(105, 95)))
fisher.test(rbind(c(109, 105), c(91, 95)))
```


```{r}





sim_data$pol_not_taken = n_stations - sim_data$pol_taken
sim_data$psd_not_taken = n_stations - sim_data$psd_taken
sim_data$pol_rate = sim_data$pol_taken / n_stations
sim_data$psd_rate = sim_data$psd_taken / n_stations

sim_i = 1

fisher.test(matrix(sim_data[i, ]))

sim_data_i = 
  matrix(
    sim_data[sim_i, 1:4],
    nrow = 2,
    byrow = TRUE,
           dimnames = list(
             pol = c("taken", "not_taken"),
             psd = c("taken", "not_taken"))
)

TeaTasting <-
  matrix(c(3, 1, 1, 3),
       nrow = 2,
       dimnames = list(Guess = c("Milk", "Tea"),
                       Truth = c("Milk", "Tea")))

TeaTasting
sim_data_i


fisher.test(
  x = factor(rbinom(n_stations, 1, pred_rate), labels = c("taken", "not_taken")),
factor(rbinom(n_stations, 1, pred_rate), labels = c("taken", "not_taken")))

fisher.test(TeaTasting)
fisher.test(sim_data_i)
fisher.test(
  matrix(sim_data[sim_i, 1:4], nrow = 2, byrow = TRUE)
)
head(sim_data)

```




- It is intuitive: if we find good *evidence* that the predation rates are different, it's reasonable to conclude there is a  *true difference*.
- It's simple
- It's easy to test!




### The setup: seeds



The exp






```{r}
seed_dat = 
  data.frame(
    species   = c("POL", "PSD"),
    predation = c(   26,    25),
    no_pred   = c(  184,    76)
  )
fisher.test()




```

