---
title: "lab 1 Q2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


Run the following two lines of code in the console and consider the differences:

```{r, eval = FALSE, echo=TRUE}
c_1 = c(1, 2, 3)
c_2 = "c(1, 2, 3)"
```

Answer the following:

- Q1 (1pt). Is c_1 a variable, or a function?
- Q2 (1pt). Is c_2 a variable, or a function?
- Q3 (1pt). If c_1 and c_2 have different values, why?


Solution
========

- Both are variables
- c_1 is a vector
- c_2 is a string


Meta-information
================
extitle: lab 1 Q2
extype: string
exsolution: nil
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 8
expts: 3