---
title: "Lab 1 Q3: matrices 1"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

Instructions

1. Create a numeric vector of length 3 called `my_vec`.  It should contain the integers from 1 to 3.
2. Build a matrix using the following code:
```{r eval = FALSE, echo = TRUE}
mat_1 = matrix(my_vec)
```

Q1 (1pt.):  What are the dimensions of the matrix (i.e. how many rows and columns)?

Q2 (2pts.):  Write R code to retrieve the element of `mat_1` that has a value of `3`.

Paste your answer and R code into the text entry box below.




Solution
========

Q1: Ideally they would use dim() to find that the matrix has 3 rows, 1 column.
Q2: Multiple possible answers including:

```{r eval = FALSE}
my_vec = 1:3
mat_1 = matrix(my_vec)

dim(mat_1)

mat_1[mat_1 == 3]

mat_1[which(mat_1 == 3)]

```




Meta-information
================
extitle: Lab 1 Q3: matrices 1
extype: string
exsolution: nil
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expts: 3