---
title: "Lab 1 Q4: matrices 2"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


```{r, echo = FALSE, results = "hide"} 
source(here::here("formatting", "functions", "rmd_functions.R"))

# include_supplement(
#     "",
#   dir = here::here(),
  # recursive = TRUE)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

You will use `my_vec` from the previous question again.

Instructions:

1. Create a matrix `mat_2` that has two rows and three columns using `my_vec`.  Do not use the `c()` or `rep()` functions.

2. Create a matrix `mat_3` that has three rows and two columns using `my_vec`.  Do not use the `c()` or `rep()` functions.



- Q1 (1pt.): Paste the code you used to create `mat_2`.
- Q2 (1pt.): Paste the code you used to create `mat_3`.
- Q3 (1pt.): Did R use **rows** or **columns** to recycle the values in `my_vec`?
- Q4 (1pt.): Create a matrix, `mat_4`, with a number of elements that is not a multiple of 3 and paste the code into the editor.
- Q5 (1pt.): How did R handle the recycling of values of `my_vec` in `mat_4`?



Solution
========

```{r eval = FALSE, echo = TRUE}
mat_2 = matrix(data = my_vec, nrow = 2, ncol = 3)
mat_3 = matrix(data = my_vec, nrow = 3, ncol = 2)

mat_4 = matrix(data = my_vec, nrow = 2, ncol = 5)

print(mat_2)
print(mat_3)
print(mat_4)
```





Meta-information
================
extitle: Lab 1 Q4: matrices 2
extype: string
exsolution: nil
exname: R Introduction
exsection: R Introduction
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 5
expts: 5