---
title: "catastrophic histogram"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```

<div class="instructionsborder">

### Instructions

- Create a histogram of the salamander reproduction catastrophic rates.
  - Make sure you include an appropriate title and label for the x-axis.
- Save your plot to an image file.

</div>


<div class="questionborder">

### Q 1 pt (file upload)

- Upload your image.

</div>




Solution
========

My implementation:

```{r reuslts='hold'}
catrate = read.csv(here("data", "catrate.csv"))
hist(
  catrate$cat.rate, 
  xlab = "Catastrophe Rate", 
  main = "Histogram of Catastrophe Rates")
```



Meta-information
================
extitle: catastrophic histogram

extype: string

exsolution: nil

expoints: 1

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE

exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 15

exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE
