---
title: "catastrophic normality test"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
require(here)
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructionsborder">

### Instructions

Conduct a Shapiro-Wilk test of normality of the salamander catastrophic rates.


</div>


<div class="questionborder">

### Q 1 pt (R code)

- Paste the code you used to conduct the test in the text box.

</div>




Solution
========

My implementation:

```{r answer_key, fig.width=4, fig.height=3}
catrate = read.csv(here::here("data", "catrate.csv"))
hist(
  catrate$cat.rate, 
  xlab = "Catastrophe Rate", 
  main = "Histogram of Catastrophe Rates")
```



Meta-information
================

extitle: catastrophic normality test

extype: string

exsolution: nil

expoints: 1

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

