---
title: "catastrophic normality results"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(here)

```





<div class="instructions">

### Instructions

Conduct a Shapiro-Wilk test of normality of the salamander catastrophic rates and interpret the output.


</div>


<div class="questions">

### Q1 1 pt (short answer)

- What is the null hypothesis for the Shapiro test?

### Q2 1 pt (short answer)

- Based on the Shapiro test results, is there strong evidence that the sample came from a non-normally-distributed population?

</div>




Solution
========

My implementation:

```{r reuslts='hold'}
catrate = read.csv(here("data", "catrate.csv"))
hist(
  catrate$cat.rate, 
  xlab = "Catastrophe Rate", 
  main = "Histogram of Catastrophe Rates")
```



Meta-information
================
extitle: catastrophic normality results

extype: string

exsolution: nil

expoints: 2

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15


