---
title: "one-sample t-test interpretation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
require(here)
```


<div class="instructions">

### Instructions

Interpret the results of your one-sample t-test of the alternative hypothesis that the catastrophic rate is different from the pond late-filling rate.  

</div>


<div class="question">

### Q1 1 pt (numeric)

- What is the p-value from your t-test?

### Q2 1 pt (numeric, short answer)

- What is the confidence interval for the difference between the null hypothesis and alternative hypothesis means?  Did it include zero?

### Q3 2 pt (short answer/essay)

- Considering the results from your t-test, did you conclude that there was strong evidence to reject the null hypothesis?
  - Make sure you justify your answer using the output of the t-test.

</div>




Solution
========


T t-test has a low p-value

```{r reuslts='hold'}
catrate = read.csv(here("data", "catrate.csv"))
t.test(catrate$cat.rate, mu = 2/7)
```





Meta-information
================
extitle: one-sample t-test interpretation

extype: string

exsolution: nil

expoints: 4

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15



