---
title: "one-sample rank sum test"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    includes:
      after_body: !expr here::here("formatting", "js", "rmd.js")

---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="instructionsborder">

### Instructions

Conduct a one-sample Wilcoxon rank sum test of the alternative hypothesis that the catastrophic rate is different from the pond late-filling rate.  

- Review the assignment walkthrough if needed.

</div>


<div class="questionborder">

### Q1 1 pt (R code)

- Paste the code you used to conduct the test.
  - Hint: your answer should be only one line of code.

</div>




Solution
========

```{r reuslts='hold', eval=FALSE}
catrate = read.csv(here::here("data", "catrate.csv"))
wilcox.test(catrate$cat.rate, mu = 2/7)
```



Meta-information
================
extitle: one-sample rank sum test

extype: string

exsolution: nil

expoints: 1

exname: individual_week_08
exsection: individual_week_08
exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

