---
title: "one-sample rank sum test interpretation"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="instructionsborder">

### Instructions

Interpret the results of your one-sample Wilcoxon rank sum test of the alternative hypothesis that the catastrophic rate is different from the pond late-filling rate.  

</div>


<div class="questionborder">

### Q1 1 pt (numeric)

- What is the p-value from your test?

### Q2 2 pt (short answer/essay)

- Considering the results from your rank sum test, did you conclude that there was strong evidence to reject the null hypothesis?
  - Make sure you justify your answer using the output of the test.

</div>




Solution
========

Q1: p-value = 0.006275

Q2: yes, test has p < 0.05


```{r}
catrate = read.csv(here::here("data", "catrate.csv"))
wilcox.test(catrate$cat.rate, mu = 2/7)
boxplot(catrate$cat.rate)
```





Meta-information
================
extitle: one-sample rank test interpretation

extype: string

exsolution: nil

expoints: 3

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

