---
title: "flipper normality tests"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="instructionsborder">

### Instructions

Conduct Shapiro-Wilk tests of normality for the flipper lengths of Chinstrap and Adelie penguins.


</div>


<div class="questionborder">

### Q1 1 pt (R code)

- Paste the code you used to conduct the test in the text box.

### Q2 1 pt (short answer)

- Were the flipper lengths normally-distributed for each species?


</div>




Solution
========

Q1: 

- `shapiro.test(dat_adelie$flipper_length_mm)`
- `shapiro.test(dat_chinstrap$flipper_length_mm)`

<div class="toggle"><button>Click to show/hide data import code</button>
<br>

```{r ans_key, eval = FALSE}
require(palmerpenguins)
dat_adelie    = subset(penguin_dat, species == "Adelie")
dat_chinstrap = subset(penguin_dat, species == "Chinstrap")
shapiro.test(dat_adelie$flipper_length_mm)
shapiro.test(dat_chinstrap$flipper_length_mm)
```
</div>

Q2: yes, both p-values were high.

Meta-information
================
extitle: flipper normality tests

extype: string

exsolution: nil

expoints: 2

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

