---
title: "flipper histograms"
output:  
  html_document:
    theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    # includes:
      # after_body: !expr here::here("formatting", "js", "rmd.js")
editor_options: 
  chunk_output_type: console
---

```{r load_data, echo = FALSE, results='hide'}
require(palmerpenguins)
dat_adelie    = subset(penguins, species == "Adelie")
dat_chinstrap = subset(penguins, species == "Chinstrap")
```

```{r plot, eval = FALSE, echo = FALSE}
par(mfrow = c(1, 2))
hist(dat_adelie$flipper_length_mm)
hist(dat_chinstrap$flipper_length_mm)
```


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```


<div class="instructionsborder">

### Instructions

Create a single figure consisting of histograms of flipper lengths of Adelie and Chinstrap penguins.

Your figure should have two histograms arranged in one row.

Save your figure to a .png file.  Your image file needs to have appropriate dimensions such that the two histograms are not vertically stretched.

- Hint: Check out the `width` and `height` arguments to `png()`.
- Hint: Remember the `par()` function?  Which argument did we use to include multiple plots in the same figure?

</div>


<div class="questionborder">

### Q1 2 pt (file upload)

- Upload your image file.

</div>




Solution
========



```{r plot, ref.label = "answer_plot", fig.asp = 1/1.5, echo = FALSE}
```

<div class="toggle"><button>Click to show/hide plot code</button>
<br>
```{r ref.label="answer_plot", eval = FALSE, echo = TRUE}
```
</div>


<script>
  $(".toggle").click(function() {
    $(this).toggleClass("open");
  });
</script>

Meta-information
================
extitle: flipper histograms

extype: string

exsolution: nil

expoints: 2

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: noinline
exextra[essay_required,logical]: FALSE
exextra[essay_fieldlines,numeric]: 1
exextra[essay_attachments,numeric]: 1
exextra[essay_attachmentsrequired,logical]: TRUE

