---
title: "two-sample t-test"
output:  
  html_document:
    # theme: readable
    css: !expr here::here("formatting", "css", "styles.css")
    toc: TRUE
    toc_float: TRUE
    includes:
      after_body: !expr here::here("formatting", "js", "rmd2.js")
---


Question
========

```{r CSS, echo = FALSE, results = "asis"}
cat(readLines(here::here("formatting", "css", "styles.css")))
```





<div class="instructionsborder">

### Instructions

Conduct a two-sample t-test of the alternative hypothesis that the Adelie penguins have different flipper lengths than Chinstrap penguins.  


</div>


<div class="questionborder">

### Q1 2 pt (short answer/essay)

- State the alternative hypothesis of the test, in plain nontechnical English.
  - Consider whether you used a one- or two- tailed test.

### Q2 1 pt (R code)

- Paste the code you used to conduct the t-test.
  - Hint: your answer should only be a single line of code.


</div>




Solution
========

Q1: Flipper length differs for Chinstrap and Adelie penguins.

Q2: `t.test(catrate$cat.rate, mu = 2/7)`

<button onclick="myTog()">Show/hide code</button>

<div id="toggleDIV">

```{r reuslts='hold', eval = FALSE}
catrate = read.csv(here::here("data", "catrate.csv"))
t.test(catrate$cat.rate, mu = 2/7)
```
</div>



Meta-information
================
extitle: two-sample t-test

extype: string

exsolution: nil

expoints: 3

exname: individual_week_08
exsection: individual_week_08

exextra[essay,logical]: TRUE
exextra[essay_format,character]: editor
exextra[essay_required,logical]: TRUE
exextra[essay_fieldlines,numeric]: 15

