---
title: "in-class tuesday Oct 20"
author: "Mike Nelson"
date: "10/20/2020"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Here is a demonstration of how to use R Markdown!




## Making a new code chunk

The shortcut key (on windows) is Ctrl-Alt-i

On Mac: Command-Option-i

If I press this key combo, RStudio will insert a new code chunk!


If I want, I can give the chunk a name and it will appear in the navigation pane.

```{r how to make a new chunk}


```





## Reading a csv file

I'm going to read some data from a csv (comma-separated-values) file.

I can use read.csv() along with the path to the file to read the data into a data.frame.

Here's an example:



```{r read_data}

require(here)

dat_birds = read.csv(here("data", "hab.sta.csv"))
dat_birds <- read.csv(here("data", "hab.sta.csv"))

# Here's a comment


```



## Previewing the contents of a data.frame

I can use the head() function to print the first few lines of a data.frame.
Here's an example with the birds data


```{r head_function_demo}
head(dat_birds)

```


## Demo plot


I can use hist() to create a histogram:


When I plot aspect, the last column is weird...

```{r simple plot}

hist(dat_birds$aspect)

```


## How to customize a histogram


I can title using the argument main = 




```{r custom histogram 1}

hist(
  dat_birds$aspect, 
  main = "Histogram of study site aspect",
  breaks = 36)



```


